---
title: "Impact of read length on alternative splicing analysis (U2AF2 KD)"
author: "Annika Ladwig"
output:
  prettydoc::html_pretty:
    theme: architect
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE,
                      results=TRUE, crop=NULL)
```

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(ggpubr)
library("RColorBrewer")
library("ggVennDiagram")
library("wesanderson")
library("colorspace")
library(UpSetR)
library(ggforce)
library(ggpointdensity)
library(viridis)
```

```{r}
# parameters
probabilityThreshold = 0.9

# theme
source("theme_paper.R")
theme_set(theme_paper())
```

```{r}
###################
# defined functions
###################

# function to extract relevant rows and columns
createTable <- function(file) {
  file = file[-1] %>%
    slice(7, 24, 27, 29, 31) %>%
    rename(n = 1)
  file = cbind(Reads = c("01 uniquely mapped", "02 multiple mapped", 
                         "03 unmapped 1", "04 unmapped 2", "05 unmapped 3"), file)
  file[nrow(file) + 1,] = c("03 unmapped", sum(c(as.numeric(file[3:5, 2]))))
  file = file %>% slice(1, 2, 6)
}

# function to load files as data frames
loadFiles <- function(x){
  x = read.table(file = x,
                 header = TRUE,
                 sep = "\t",
                 stringsAsFactors = FALSE)
}

# function to count semicolons of column "mean_dpsi_per_lsv_junction"
# to find out the number of junctions per LSV
# & store this information in new column "nJunctions"
countJunctions <- function(x) {
  SRSF6_junctionsPerLSV = x$mean_dpsi_per_lsv_junction %>%
   str_count(pattern = ";") + 1
  x$nJunctions = SRSF6_junctionsPerLSV
  x
}

# count the amount of LSVs with 2, 3, 4, ... junctions
# create data frame with this information
complexityLSVs <- function(x) {
   junctionsPerLSV = x$nJunctions %>%
    table(dnn="junctionsPerLSV") %>%
    as.data.frame
   junctionsPerLSV
}

# function to extract max. dPSI for each LSV and add new column with this information
extractMaxDPSIs <- function(x) {
  maxDeltaPSIs = x$mean_dpsi_per_lsv_junction %>%
    str_split(pattern = ";") %>%
    sapply(FUN = function(dPSIs){
      return(
        dPSIs %>% as.numeric %>% abs %>% max
      )
    })
  x$maxDeltaPSI = maxDeltaPSIs
  x
}

# function to extract max. probability for each LSV and add new column with this information
extractMaxProbs <- function(x) {
  maxProbs = x$probability_changing %>%
    str_split(pattern=";") %>%
    sapply(FUN=function(probs){
      return(
        probs %>% as.numeric %>% max
      )
    })
  x$maxProb = maxProbs
  x
}
```

# 1 - Influence on read alignment (Fig 1D)

The influence of read length on read alignment is determined by plotting the fraction of unique, multiple and unmapped reads for each read length (2 replicates each for knockdown and control).The alignment statistics are obtained from STAR. 

```{r}
# load star log files 
StarDir = "STAR_Log.final.out"
StarLogFiles=list.files(path=StarDir)
setwd(StarDir)
StarLogFiles = lapply(StarLogFiles, read.delim)

# number of input reads for each replicate
for(file in seq(1, 16, 4)){
  print(StarLogFiles[[file]][4,2])
}

# create table with list of star log files
StarLogFiles = lapply(StarLogFiles, createTable)

# add new column with read length to the tables to differ the data
# when merging them into a common table 
for(i in seq(1, 16, 4)){
  StarLogFiles[[i]]$readLength = "100"
}
for(i in seq(2, 16, 4)){
  StarLogFiles[[i]]$readLength = "075"
}
for(i in seq(3, 16, 4)){
  StarLogFiles[[i]]$readLength = "050"
}
for(i in seq(4, 16, 4)){
  StarLogFiles[[i]]$readLength = "025"
}

# one table for each experiment for different read lengths
StarLogTable_1 = rbind(StarLogFiles[[1]], StarLogFiles[[2]], StarLogFiles[[3]], StarLogFiles[[4]])
StarLogTable_2 = rbind(StarLogFiles[[5]], StarLogFiles[[6]], StarLogFiles[[7]], StarLogFiles[[8]])
StarLogTable_3 = rbind(StarLogFiles[[9]], StarLogFiles[[10]], StarLogFiles[[11]], StarLogFiles[[12]])
StarLogTable_4 = rbind(StarLogFiles[[13]], StarLogFiles[[14]], StarLogFiles[[15]], StarLogFiles[[16]])

# star log output as stacked bar chart for each experiment
StarReport_StackedBarChart = ggarrange(
  ggplot(StarLogTable_1, aes(x = readLength, y = as.numeric(n), fill = Reads)) +
    geom_col(position = position_fill(reverse = TRUE)) +
    xlab("read length [nt]") +
    ylab("% reads") +
    scale_x_discrete(labels = c("36", "50", "75", "100")) +
    ggtitle("knockdown 1") +
    theme(legend.title = element_blank()) +
    scale_fill_manual(values = wes_palette("GrandBudapest1")),
  
  ggplot(StarLogTable_2, aes(x = readLength, y = as.numeric(n), fill = Reads)) +
    geom_col(position = position_fill(reverse = TRUE)) +
    xlab("read length [nt]") +
    ylab("% reads") +
    scale_x_discrete(labels = c("36", "50", "75", "100")) +
    ggtitle("control 1") +
    theme(legend.title = element_blank()) +
    scale_fill_manual(values = wes_palette("GrandBudapest1")),
  
  ggplot(StarLogTable_3, aes(x = readLength, y = as.numeric(n), fill = Reads)) +
    geom_col(position = position_fill(reverse = TRUE)) +
    xlab("read length [nt]") +
    ylab("% reads") +
    scale_x_discrete(labels = c("36", "50", "75", "100")) +
    ggtitle("knockdown 2") +
    theme(legend.title = element_blank()) +
    scale_fill_manual(values = wes_palette("GrandBudapest1")),
  
  ggplot(StarLogTable_4, aes(x = readLength, y = as.numeric(n), fill = Reads)) +
    geom_col(position = position_fill(reverse = TRUE)) +
    xlab("read length [nt]") +
    ylab("% reads") +
    scale_x_discrete(labels = c("36", "50", "75", "100")) +
    ggtitle("control 2") +
    theme(legend.title = element_blank()) +
    scale_fill_manual(values = wes_palette("GrandBudapest1")),
  
  common.legend = TRUE, legend = "bottom",
  ncol = 4, nrow = 1
  )

StarReport_StackedBarChart = annotate_figure(StarReport_StackedBarChart, top = text_grob("Influence on read alignment", size = 12, just = "centre", face = "bold"))
StarReport_StackedBarChart
```

# 2.1 - Influence on LSV detection (Fig 2A)

The influence of read length on LSV detection is determined by counting the LSV hits (rows) in the tsv files received from MAJIQ for each read length. 

```{r}
# load dPSI tsv files from MAJIQ Quantifier as list
dpsiDir = "MAJIQ/read_length"
dpsiFiles=list.files(path = dpsiDir, pattern = ".tsv")
setwd(dpsiDir)

# load tsv files as data frames
dPSIs = lapply(dpsiFiles, loadFiles)
```

```{r}
# count detected LSVs
nLSVs_readLength_36 = nrow(dPSIs[[1]])
nLSVs_readLength_50 = nrow(dPSIs[[2]])
nLSVs_readLength_75 = nrow(dPSIs[[3]])
nLSVs_readLength_100 = nrow(dPSIs[[4]])

# create dataframe with nLSVs 
nLSVs_df = data.frame(readLength=c("036", "050", "075", "100"),
                      nLSVs=c(nLSVs_readLength_36, nLSVs_readLength_50, 
                              nLSVs_readLength_75, nLSVs_readLength_100))

# number of LSVs bar chart
nLSVs_bar = ggplot(data=nLSVs_df, aes(x = readLength, y = nLSVs, fill = readLength)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  ylim(0, 50500) +
  xlab("read length [nt]") +
  ylab("# LSVs") +
  ggtitle("A") +
  scale_fill_manual(name = "read length [nt]", 
                    values = c("#FAD8B4", "#F7B56D", "#DE872A", "#C85813"),  
                    labels = c("36", "50", "75", "100")) +
  scale_x_discrete(labels = c("36", "50", "75", "100")) +
  geom_text(aes(label=nLSVs), position=position_dodge(width=0.9), vjust=-0.25) 
```

# 2.2 - Influence on LSV complexity (Fig 2B)

The LSV complexity describes how many junctions an LSV has. To measure the influence on this, the LSVs with 2, 3, 4 or more than 4 junctions are counted for each read length. 

```{r}
# count junctions of each LSV
# create new list with the data frames that contain the information about the complexity of the LSVs
dPSIs = lapply(dPSIs, countJunctions)
complexityLSVs = lapply(dPSIs, complexityLSVs)

# add new column with read length to differ the data 
# when merging the single data frames to one 
complexityLSVs[[1]]$readLength = "036"
complexityLSVs[[2]]$readLength = "050"
complexityLSVs[[3]]$readLength = "075"
complexityLSVs[[4]]$readLength = "100"

# create data frame out of all read length to build a stacked bar chart
nJunctions_df <- rbind(complexityLSVs[[1]], complexityLSVs[[2]], complexityLSVs[[3]], complexityLSVs[[4]])

# all LSVs with junctions > 4 are grouped into one category
# -> all LSVs with > 4 junctions get the same color in following stacked bar chart
for (i in 1:nrow(nJunctions_df)){
  if (as.integer(nJunctions_df[i, "junctionsPerLSV"]) >= 5) {
    nJunctions_df[i, "junctionsPerLSV"] = 5
  }
}

# stacked bar chart to show complexity of LSVs for each read length
StackedBarChart_nJunctions = ggplot(nJunctions_df, aes(x = readLength, y = Freq)) +
  geom_bar(stat = "identity", aes(fill = readLength), position = "fill", show.legend = FALSE) +
  scale_fill_manual(values = c("#FAD8B4", "#F7B56D", "#DE872A", "#C85813"),
                    labels = c("36", "50", "75", "100"),
                    name = "read length [nt]") + 
  geom_bar(stat = "identity", fill = "black", aes(alpha = junctionsPerLSV), position = "fill", show.legend = FALSE) +
  scale_alpha_manual(values = c(0, 0.3, 0.55, 0.8),
                     name = "junctions per LSV", 
                     breaks = c("2", "3", "4", "5"),
                     labels = c("2", "3", "4", "> 4")) +
  scale_x_discrete(labels = c("36", "50", "75", "100")) +
  xlab("read length [nt]") +
  ylab("fraction LSVs [%]") +
  ggtitle("Influence on LSV complexity") 
```

# 3 - Influence on the significance of LSVs (Fig 3)

Influence of read length on the significance of LSVs is determined by plotting the fraction of significantly regulated LSVs detected for each read length. LSVs are considered significant if dPSI is at least 0.05 and probability changing is at least 0.9. 

For each LSV, the maximum dpsi and max probability changing is chosen. These usually belong to the same junction.

```{r}
# extract max. dPSI for each LSV and add new column with this information
dPSIs = lapply(dPSIs, extractMaxDPSIs)

# extract max. probability for each LSV and add a new column with this information
dPSIs = lapply(dPSIs, extractMaxProbs)
```

## dPSI against probability changing (Fig 3A)
```{r}
# plot max. dPSI against max. probability for each LSV (100 nt long reads)
scatter_readLength_100 = ggplot(dPSIs[[4]], aes(x = maxDeltaPSI, y = maxProb)) + geom_pointdensity() +
                         scale_color_viridis(option = "rocket") +
                         geom_hline(yintercept = probabilityThreshold, col = "red3", linetype = "dashed", size = 0.62) +
                         geom_vline(xintercept = 0.05, col = "red3", linetype = "dashed", size = 0.62) +
                         xlab("\u0394\u03A8") 
```

## Number and fraction of detected significant LSVs (Fig 3B)
```{r}
# add new column "significant" for each read length which could be TRUE/FALSE 
# TRUE if maxProb >= 0.9 and maxDPSI >= 0.05 (significance threshold)
dPSIs[[1]]$significant = dPSIs[[1]]$maxProb >= probabilityThreshold & dPSIs[[1]]$maxDeltaPSI >= 0.05
dPSIs[[2]]$significant = dPSIs[[2]]$maxProb >= probabilityThreshold & dPSIs[[2]]$maxDeltaPSI >= 0.05
dPSIs[[3]]$significant = dPSIs[[3]]$maxProb >= probabilityThreshold & dPSIs[[3]]$maxDeltaPSI >= 0.05
dPSIs[[4]]$significant = dPSIs[[4]]$maxProb >= probabilityThreshold & dPSIs[[4]]$maxDeltaPSI >= 0.05

# significant LSVs for each read length
print("# significant LSVs")
table(dPSIs[[1]]$significant)
table(dPSIs[[2]]$significant)
table(dPSIs[[3]]$significant)
table(dPSIs[[4]]$significant)

# count significant and not significant LSVs for each read length
nSignificant_36 = sum(dPSIs[[1]]$significant)
nNotSignificant_36 = (nLSVs_readLength_36-sum(dPSIs[[1]]$significant))
nSignificant_50 = sum(dPSIs[[2]]$significant)
nNotSignificant_50 = (nLSVs_readLength_50-sum(dPSIs[[2]]$significant))
nSignificant_75 = sum(dPSIs[[3]]$significant)
nNotSignificant_75 = (nLSVs_readLength_75-sum(dPSIs[[3]]$significant))
nSignificant_100 = sum(dPSIs[[4]]$significant)
nNotSignificant_100 = (nLSVs_readLength_100-sum(dPSIs[[4]]$significant))

#############################
# number of significant LSVs
#############################

# dataframe with number of significantly regulated LSVs
nSignificantLSVs_df_absolute = data.frame(readLength=c("036", "050", "075", "100"),
                      nSignificantLSVs=c(nSignificant_36, nSignificant_50, nSignificant_75, nSignificant_100))

# number of signifciant LSVs bar chart
nSignificantLSVs_bar_absolute = 
  ggplot(data=nSignificantLSVs_df_absolute, aes(x = readLength, y = nSignificantLSVs, fill = readLength)) +
  geom_bar(stat = "identity") +
  xlab("read length [nt]") +
  ylab("# significant LSVs") +
  scale_fill_manual(name = "read length [nt]", values = c("#FAD8B4", "#F7B56D", "#DE872A", "#C85813"), 
                    labels = c("36", "50", "75", "100")) +
  scale_x_discrete(labels = c("36", "50", "75", "100")) +
  geom_text(aes(label=nSignificantLSVs), position=position_dodge(width=0.9), vjust=1) 

###############################
# fraction significant LSVs
###############################

# create data frame with a column for the read length
# and a column indicating whether the associated LSV is significant or not 
significantLSVs_df_relative = data.frame(comparison=c(rep("036", nrow(dPSIs[[1]])),
                                                   rep("050", nrow(dPSIs[[2]])), 
                                                   rep("075", nrow(dPSIs[[3]])), 
                                                   rep("100", nrow(dPSIs[[4]]))),
                                      significant=c(dPSIs[[1]]$significant,
                                                    dPSIs[[2]]$significant, 
                                                    dPSIs[[3]]$significant, 
                                                    dPSIs[[4]]$significant))

# differ between the significance of LSVs of different sample combinations to plot them in different colors 
for (i in 1:nrow(significantLSVs_df_relative)){
  if (significantLSVs_df_relative[i, "comparison"] == "036" && significantLSVs_df_relative[i, "significant"] == TRUE){
    significantLSVs_df_relative[i, "significant"] = "TRUE read length 036"
  }
  if (significantLSVs_df_relative[i, "comparison"] == "050" && significantLSVs_df_relative[i, "significant"] == TRUE){
    significantLSVs_df_relative[i, "significant"] = "TRUE read length 050"
  }
  if (significantLSVs_df_relative[i, "comparison"] == "075" && significantLSVs_df_relative[i, "significant"] == TRUE){
    significantLSVs_df_relative[i, "significant"] = "TRUE read length 075"
  }
  if (significantLSVs_df_relative[i, "comparison"] == "100" && significantLSVs_df_relative[i, "significant"] == TRUE){
    significantLSVs_df_relative[i, "significant"] = "TRUE read length 100"
  }
}

# stacked bar chart showing fraction of significant LSVs among all LSVs for each read length   
nSignificantLSVs_bar_relative = ggplot(data=significantLSVs_df_relative, aes(x = comparison, fill = factor(significant))) +
  geom_bar(position = "fill", show.legend = FALSE) +
  xlab("read length [nt]") +
  ylab("fraction LSVs [%]") +
  coord_cartesian(ylim = c(0, 0.07)) +
  scale_fill_manual(name= "significant", values = c("lightgrey", "#FAD8B4", "#F7B56D", "#DE872A", "#C85813"), labels = c("FALSE", "TRUE", "TRUE", "TRUE", "TRUE")) +
  scale_x_discrete(labels = c("36", "50", "75", "100")) +
  scale_color_manual(name = "mean 2 samples") 
```

## Shared significant LSVs between different datasets (Fig 3C)
```{r}
# check number of regulated LSVs
# TRUE = 1, FALSE = 0
sum(dPSIs[[1]]$significant)
sum(dPSIs[[2]]$significant)
sum(dPSIs[[3]]$significant)
sum(dPSIs[[4]]$significant)

# extract all significant LSVs for each read length
significantLSVs_36 = dPSIs[[1]]$lsv_id[dPSIs[[1]]$significant == TRUE]
significantLSVs_50 = dPSIs[[2]]$lsv_id[dPSIs[[2]]$significant == TRUE]
significantLSVs_75 = dPSIs[[3]]$lsv_id[dPSIs[[3]]$significant == TRUE]
significantLSVs_100 = dPSIs[[4]]$lsv_id[dPSIs[[4]]$significant == TRUE]

# list with all significant LSVs for each read length
significantLSVs_all = list("36" = c(significantLSVs_36), "50" = c(significantLSVs_50), 
                           "75" = c(significantLSVs_75), "100" = c(significantLSVs_100))

# upset plot
pdf(file=".../plots/upset_read_length.pdf", width=3.6, height=3) 
upset(fromList(significantLSVs_all), 
      order.by = "freq", 
      sets.bar.color = c("#FAD8B4", "#F7B56D", "#DE872A", "#C85813"), 
      sets.x.label = "# significant LSVs", 
      mainbar.y.label = "intersection")
dev.off()
```


# Influence on size of detected splicing changes (Fig 4)

Influence on the size of detected splicing changes is determined by plotting the cumulative fraction of dPSIs and probability changing values for each read length. 

```{r}
dPSIs[[1]]$readLength = "036"
dPSIs[[2]]$readLength = "50"
dPSIs[[3]]$readLength = "75"
dPSIs[[4]]$readLength = "100"

# cumulative fraction of dPSIs
ecdf_deltaPSI = ggplot() +
  stat_ecdf(aes(x = maxDeltaPSI), color = "black", data = dPSIs[[1]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "036"), data = dPSIs[[1]], size = 1, pad = FALSE) +
  
  stat_ecdf(aes(x = maxDeltaPSI), color = "black", data = dPSIs[[2]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "050"), data = dPSIs[[2]], size = 1, pad = FALSE) +
  
  stat_ecdf(aes(x = maxDeltaPSI), color = "black", data = dPSIs[[3]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "075"), data = dPSIs[[3]], size = 1, pad = FALSE) +
  
  stat_ecdf(aes(x = maxDeltaPSI), color = "black", data = dPSIs[[4]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "100"), data = dPSIs[[4]], size = 1, pad = FALSE) +
  scale_color_manual(values = c("#FAD8B4", "#FFB565", "#DE872A", "#C85813"), 
                     name = "read length  [nt]", 
                     labels = c("36", "50", "75", "100", "36", "50", "75", "100")) + 
  geom_vline(xintercept = 0.05, col = "red3", linetype = "dashed", size = 0.5) +
  coord_cartesian(xlim=c(0, 0.5)) +
  ylab("cumulative fraction") +
  xlab("dPSI") +
  theme(legend.position="none")

# cumulative fraction of probability changing values 
ecdf_prob_changing = ggplot() +
  stat_ecdf(aes(x = maxProb), color = "black", data = dPSIs[[1]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxProb, color = readLength), data = dPSIs[[1]], size = 1, pad = FALSE) +
  
  stat_ecdf(aes(x = maxProb), color = "black", data = dPSIs[[2]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxProb, color = readLength), data = dPSIs[[2]], size = 1, pad = FALSE) +
  
  stat_ecdf(aes(x = maxProb), color = "black", data = dPSIs[[3]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxProb, color = readLength), data = dPSIs[[3]], size = 1, pad = FALSE) +
  
  stat_ecdf(aes(x = maxProb), color = "black", data = dPSIs[[4]], size = 1.25, pad = FALSE) +
  stat_ecdf(aes(x = maxProb, color = readLength), data = dPSIs[[4]], size = 1, pad = FALSE) +
  scale_color_manual(name = "read length  [nt]", values = c("#FAD8B4", "#FFB565", "#DE872A", "#C85813"), 
                     labels = c("36", "50", "75", "100")) + 
  geom_vline(xintercept = 0.9, col = "red3", linetype = "dashed", size = 0.5) +
  xlim(0, 1) +
  ylab("cumulative fraction") +
  xlab("P(|dPSI| > 0.05)") +
  theme(legend.position="none")
```


## Shared binary regulation between the different datasets

Only binary LSVs are considered to ensure that the dPSI value of the same junction is taken from both data sets. 

```{r}
#################
# all binary LSVs
#################

# filter the data frames for LSVs with 2 Junctions
# significance shown
U2AF2_100_filtered = dPSIs[[4]][dPSIs[[4]]$nJunctions == 2,]
U2AF2_75_filtered = dPSIs[[3]][dPSIs[[3]]$nJunctions == 2,]
U2AF2_50_filtered = dPSIs[[2]][dPSIs[[2]]$nJunctions == 2,]
U2AF2_36_filtered = dPSIs[[1]][dPSIs[[1]]$nJunctions == 2,]

# identify shared significant LSVs for each comparison
# extract LSV IDs that intersect
shared_U2AF2_100_75 = intersect(U2AF2_100_filtered$lsv_id, U2AF2_75_filtered$lsv_id) 
length(shared_U2AF2_100_75)
shared_U2AF2_100_50 = intersect(U2AF2_100_filtered$lsv_id, U2AF2_50_filtered$lsv_id) 
length(shared_U2AF2_100_50)
shared_U2AF2_100_36 = intersect(U2AF2_100_filtered$lsv_id, U2AF2_36_filtered$lsv_id)
length(shared_U2AF2_100_36)
shared_U2AF2_75_50 = intersect(U2AF2_75_filtered$lsv_id, U2AF2_50_filtered$lsv_id)
length(shared_U2AF2_75_50)
shared_U2AF2_75_36 = intersect(U2AF2_75_filtered$lsv_id, U2AF2_36_filtered$lsv_id)
length(shared_U2AF2_75_36)
shared_U2AF2_50_36 = intersect(U2AF2_50_filtered$lsv_id, U2AF2_36_filtered$lsv_id)
length(shared_U2AF2_50_36)

# extract filtered LSVs from data frames above
# compare 100 with 75
U2AF2_100_filtered_match_75 <- U2AF2_100_filtered[match(shared_U2AF2_100_75, U2AF2_100_filtered$lsv_id),]
U2AF2_75_filtered_match_100 <- U2AF2_75_filtered[match(shared_U2AF2_100_75, U2AF2_75_filtered$lsv_id),]

# compare 100 with 50
U2AF2_100_filtered_match_50 <- U2AF2_100_filtered[match(shared_U2AF2_100_50, U2AF2_100_filtered$lsv_id),]
U2AF2_50_filtered_match_100 <- U2AF2_50_filtered[match(shared_U2AF2_100_50, U2AF2_50_filtered$lsv_id),]

# compare 100 with 36
U2AF2_100_filtered_match_36 <- U2AF2_100_filtered[match(shared_U2AF2_100_36, U2AF2_100_filtered$lsv_id),]
U2AF2_36_filtered_match_100 <- U2AF2_36_filtered[match(shared_U2AF2_100_36, U2AF2_36_filtered$lsv_id),]

# compare 75 with 50
U2AF2_75_filtered_match_50 <- U2AF2_75_filtered[match(shared_U2AF2_75_50, U2AF2_75_filtered$lsv_id),]
U2AF2_50_filtered_match_75 <- U2AF2_50_filtered[match(shared_U2AF2_75_50, U2AF2_50_filtered$lsv_id),]

# compare 75 with 36
U2AF2_75_filtered_match_36 <- U2AF2_75_filtered[match(shared_U2AF2_75_36, U2AF2_75_filtered$lsv_id),]
U2AF2_36_filtered_match_75 <- U2AF2_36_filtered[match(shared_U2AF2_75_36, U2AF2_36_filtered$lsv_id),]

# compare 50 with 36
U2AF2_50_filtered_match_36 <- U2AF2_50_filtered[match(shared_U2AF2_50_36, U2AF2_50_filtered$lsv_id),]
U2AF2_36_filtered_match_50 <- U2AF2_36_filtered[match(shared_U2AF2_50_36, U2AF2_36_filtered$lsv_id),]

# extract dPSI value of the first junction for all LSVs
# deltapsi 100 vs 75 
deltaPSI_U2AF2_100_vs_75 = U2AF2_100_filtered_match_75$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_U2AF2_75_vs_100 = U2AF2_75_filtered_match_100$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 100 vs 50 
deltaPSI_U2AF2_100_vs_50 = U2AF2_100_filtered_match_50$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_U2AF2_50_vs_100 = U2AF2_50_filtered_match_100$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 100 vs 36 
deltaPSI_U2AF2_100_vs_36 = U2AF2_100_filtered_match_36$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_36_vs_100 = U2AF2_36_filtered_match_100$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 75 vs 50
deltaPSI_U2AF2_75_vs_50 = U2AF2_75_filtered_match_50$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_50_vs_75 = U2AF2_50_filtered_match_75$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 75 vs 36
deltaPSI_U2AF2_75_vs_36 = U2AF2_75_filtered_match_36$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_36_vs_75 = U2AF2_36_filtered_match_75$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 50 vs 36
deltaPSI_U2AF2_50_vs_36 = U2AF2_50_filtered_match_36$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_36_vs_50 = U2AF2_36_filtered_match_50$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric


##########################
# significant binary LSVs
##########################

# filter the data frames for significant LSVs with 2 Junctions
U2AF2_100_filtered_sig = dPSIs[[4]][dPSIs[[4]]$nJunctions == 2 & dPSIs[[4]]$significant == TRUE,]
U2AF2_75_filtered_sig = dPSIs[[3]][dPSIs[[3]]$nJunctions == 2 & dPSIs[[3]]$significant == TRUE,]
U2AF2_50_filtered_sig = dPSIs[[2]][dPSIs[[2]]$nJunctions == 2 & dPSIs[[2]]$significant == TRUE,]
U2AF2_36_filtered_sig = dPSIs[[1]][dPSIs[[1]]$nJunctions == 2 & dPSIs[[1]]$significant == TRUE,]

# identify shared significant LSVs for each comparison
shared_U2AF2_100_U2AF2_75_sig = intersect(U2AF2_100_filtered_sig$lsv_id, U2AF2_75_filtered_sig$lsv_id) 
length(shared_U2AF2_100_U2AF2_75_sig)
shared_U2AF2_100_U2AF2_50_sig = intersect(U2AF2_100_filtered_sig$lsv_id, U2AF2_50_filtered_sig$lsv_id) 
length(shared_U2AF2_100_U2AF2_50_sig)
shared_U2AF2_100_U2AF2_36_sig = intersect(U2AF2_100_filtered_sig$lsv_id, U2AF2_36_filtered_sig$lsv_id)
length(shared_U2AF2_100_U2AF2_36_sig)
shared_U2AF2_75_U2AF2_50_sig = intersect(U2AF2_75_filtered_sig$lsv_id, U2AF2_50_filtered_sig$lsv_id)
length(shared_U2AF2_75_U2AF2_50_sig)
shared_U2AF2_75_U2AF2_36_sig = intersect(U2AF2_75_filtered_sig$lsv_id, U2AF2_36_filtered_sig$lsv_id)
length(shared_U2AF2_75_U2AF2_36_sig)
shared_U2AF2_50_U2AF2_36_sig = intersect(U2AF2_50_filtered_sig$lsv_id, U2AF2_36_filtered_sig$lsv_id)
length(shared_U2AF2_50_U2AF2_36_sig)

# extract filtered LSVs from data frames above
# compare 100 with 75
U2AF2_100_filtered_match_U2AF2_75_sig <- U2AF2_100_filtered_sig[match(shared_U2AF2_100_U2AF2_75_sig, U2AF2_100_filtered_sig$lsv_id),]

U2AF2_75_filtered_match_U2AF2_100_sig <- U2AF2_75_filtered_sig[match(shared_U2AF2_100_U2AF2_75_sig, U2AF2_75_filtered_sig$lsv_id),]

# compare 100 with 50
U2AF2_100_filtered_match_U2AF2_50_sig <- U2AF2_100_filtered_sig[match(shared_U2AF2_100_U2AF2_50_sig, U2AF2_100_filtered_sig$lsv_id),]

U2AF2_50_filtered_match_U2AF2_100_sig <- U2AF2_50_filtered_sig[match(shared_U2AF2_100_U2AF2_50_sig, U2AF2_50_filtered_sig$lsv_id),]

# compare 100 with 36
U2AF2_100_filtered_match_U2AF2_36_sig <- U2AF2_100_filtered_sig[match(shared_U2AF2_100_U2AF2_36_sig, U2AF2_100_filtered_sig$lsv_id),]

U2AF2_36_filtered_match_U2AF2_100_sig <- U2AF2_36_filtered_sig[match(shared_U2AF2_100_U2AF2_36_sig, U2AF2_36_filtered_sig$lsv_id),]

# compare 75 with 50
U2AF2_75_filtered_match_U2AF2_50_sig <- U2AF2_75_filtered_sig[match(shared_U2AF2_75_U2AF2_50_sig, U2AF2_75_filtered_sig$lsv_id),]

U2AF2_50_filtered_match_U2AF2_75_sig <- U2AF2_50_filtered_sig[match(shared_U2AF2_75_U2AF2_50_sig, U2AF2_50_filtered_sig$lsv_id),]

# compare 75 with 36
U2AF2_75_filtered_match_U2AF2_36_sig <- U2AF2_75_filtered_sig[match(shared_U2AF2_75_U2AF2_36_sig, U2AF2_75_filtered_sig$lsv_id),]

U2AF2_36_filtered_match_U2AF2_75_sig <- U2AF2_36_filtered_sig[match(shared_U2AF2_75_U2AF2_36_sig, U2AF2_36_filtered_sig$lsv_id),]

# compare 50 with 36
U2AF2_50_filtered_match_U2AF2_36_sig <- U2AF2_50_filtered_sig[match(shared_U2AF2_50_U2AF2_36_sig, U2AF2_50_filtered_sig$lsv_id),]

U2AF2_36_filtered_match_U2AF2_50_sig <- U2AF2_36_filtered_sig[match(shared_U2AF2_50_U2AF2_36_sig, U2AF2_36_filtered_sig$lsv_id),]

# extract dPSI value of the first junction for all LSVs
# deltapsi 100 vs 75 
deltaPSI_U2AF2_100_vs_U2AF2_75_sig = U2AF2_100_filtered_match_U2AF2_75_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_U2AF2_75_vs_U2AF2_100_sig = U2AF2_75_filtered_match_U2AF2_100_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 100 vs 50 
deltaPSI_U2AF2_100_vs_U2AF2_50_sig = U2AF2_100_filtered_match_U2AF2_50_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_U2AF2_50_vs_U2AF2_100_sig = U2AF2_50_filtered_match_U2AF2_100_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 100 vs 36 
deltaPSI_U2AF2_100_vs_U2AF2_36_sig = U2AF2_100_filtered_match_U2AF2_36_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_36_vs_U2AF2_100_sig = U2AF2_36_filtered_match_U2AF2_100_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 75 vs 50
deltaPSI_U2AF2_75_vs_U2AF2_50_sig = U2AF2_75_filtered_match_U2AF2_50_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_50_vs_U2AF2_75_sig = U2AF2_50_filtered_match_U2AF2_75_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 75 vs 36
deltaPSI_U2AF2_75_vs_U2AF2_36_sig = U2AF2_75_filtered_match_U2AF2_36_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_36_vs_U2AF2_75_sig = U2AF2_36_filtered_match_U2AF2_75_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi 50 vs 36
deltaPSI_U2AF2_50_vs_U2AF2_36_sig = U2AF2_50_filtered_match_U2AF2_36_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric                   

deltaPSI_U2AF2_36_vs_U2AF2_50_sig = U2AF2_36_filtered_match_U2AF2_50_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric
```

```{r}
# build data frame out of vectors with dPSI values 
# create scatter plot with the data frames

# correlation data frame 100 vs 75
correlationDataFrame_100_vs_75 <- data.frame(deltaPSI_100_vs_75, deltaPSI_U2AF2_75_vs_100)
correlationDataFrame_100_vs_U2AF2_75_sig <- data.frame(deltaPSI_100_vs_U2AF2_75_sig, deltaPSI_U2AF2_75_vs_100_sig)

sharedRegulation_scatterPlot1 = ggplot(data=correlationDataFrame_100_vs_75, aes(x=deltaPSI_100_vs_75, deltaPSI_U2AF2_75_vs_100)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_100_vs_75, aes(x=deltaPSI_100_vs_75, deltaPSI_U2AF2_75_vs_100), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_100_vs_U2AF2_75_sig, aes(x=deltaPSI_100_vs_U2AF2_75_sig, y=deltaPSI_U2AF2_75_vs_100_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_100_vs_U2AF2_75_sig, aes(x=deltaPSI_100_vs_U2AF2_75_sig, y=deltaPSI_U2AF2_75_vs_100_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  coord_cartesian(xlim = c(-0.8, 0.8), ylim = c(-0.8, 0.8)) +
  coord_fixed(ratio = 1) +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  xlab("dPSI 100 nt read length") +
  ylab("dPSI 75 nt read length") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame 100 vs 50
correlationDataFrame_100_vs_50 <- data.frame(deltaPSI_100_vs_50, deltaPSI_50_vs_100)
correlationDataFrame_100_vs_50_sig <- data.frame(deltaPSI_100_vs_50_sig, deltaPSI_50_vs_100_sig)

sharedRegulation_scatterPlot2 = ggplot(data=correlationDataFrame_100_vs_50, aes(x=deltaPSI_100_vs_50, deltaPSI_50_vs_100)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_100_vs_50, aes(x=deltaPSI_100_vs_50, deltaPSI_50_vs_100), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_100_vs_50_sig, aes(x=deltaPSI_100_vs_50_sig, y=deltaPSI_50_vs_100_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_100_vs_50_sig, aes(x=deltaPSI_100_vs_50_sig, y=deltaPSI_50_vs_100_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") + 
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_cartesian(xlim = c(-0.7, 0.7), ylim = c(-0.7, 0.7)) +
  xlab("dPSI 100 nt read length") +
  ylab("dPSI 50 nt read length") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold")) +
  theme(aspect.ratio=1) +
  ggtitle("D")

# correlation data frame 100 vs 36
correlationDataFrame_100_vs_36 <- data.frame(deltaPSI_100_vs_36, deltaPSI_36_vs_100)
correlationDataFrame_100_vs_36_sig <- data.frame(deltaPSI_100_vs_36_sig, deltaPSI_36_vs_100_sig)

sharedRegulation_scatterPlot3 = ggplot(data=correlationDataFrame_100_vs_36, aes(x=deltaPSI_100_vs_36, deltaPSI_36_vs_100)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_100_vs_36, aes(x=deltaPSI_100_vs_36, deltaPSI_36_vs_100), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_100_vs_36_sig, aes(x=deltaPSI_100_vs_36_sig, y=deltaPSI_36_vs_100_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_100_vs_36_sig, aes(x=deltaPSI_100_vs_36_sig, y=deltaPSI_36_vs_100_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("dPSI 100 nt read length") +
  ylab("dPSI 50 nt read length") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame 75 vs 50
correlationDataFrame_U2AF2_75_vs_50 <- data.frame(deltaPSI_U2AF2_75_vs_50, deltaPSI_50_vs_75)
correlationDataFrame_U2AF2_75_vs_50_sig <- data.frame(deltaPSI_U2AF2_75_vs_50_sig, deltaPSI_50_vs_U2AF2_75_sig)

sharedRegulation_scatterPlot4 = ggplot(data=correlationDataFrame_U2AF2_75_vs_50, aes(x=deltaPSI_U2AF2_75_vs_50, deltaPSI_50_vs_75)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_U2AF2_75_vs_50, aes(x=deltaPSI_U2AF2_75_vs_50, deltaPSI_50_vs_75), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_U2AF2_75_vs_50_sig, aes(x=deltaPSI_U2AF2_75_vs_50_sig, y=deltaPSI_50_vs_U2AF2_75_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_U2AF2_75_vs_50_sig, aes(x=deltaPSI_U2AF2_75_vs_50_sig, y=deltaPSI_50_vs_U2AF2_75_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("dPSI 75 nt read length") +
  ylab("dPSI 50 nt read length") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame 75 vs 36
correlationDataFrame_U2AF2_75_vs_36 <- data.frame(deltaPSI_U2AF2_75_vs_36, deltaPSI_36_vs_75)
correlationDataFrame_U2AF2_75_vs_36_sig <- data.frame(deltaPSI_U2AF2_75_vs_36_sig, deltaPSI_36_vs_U2AF2_75_sig)

sharedRegulation_scatterPlot5 = ggplot(data=correlationDataFrame_U2AF2_75_vs_36, aes(x=deltaPSI_U2AF2_75_vs_36, deltaPSI_36_vs_75)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_U2AF2_75_vs_36, aes(x=deltaPSI_U2AF2_75_vs_36, deltaPSI_36_vs_75), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_U2AF2_75_vs_36_sig, aes(x=deltaPSI_U2AF2_75_vs_36_sig, y=deltaPSI_36_vs_U2AF2_75_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_U2AF2_75_vs_36_sig, aes(x=deltaPSI_U2AF2_75_vs_36_sig, y=deltaPSI_36_vs_U2AF2_75_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("dPSI 75 nt read length") +
  ylab("dPSI 50 nt read length") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame 50 vs 36
correlationDataFrame_50_vs_36 <- data.frame(deltaPSI_50_vs_36, deltaPSI_36_vs_50)
correlationDataFrame_50_vs_36_sig <- data.frame(deltaPSI_50_vs_36_sig, deltaPSI_36_vs_50_sig)

sharedRegulation_scatterPlot6 = ggplot(data=correlationDataFrame_50_vs_36, aes(x=deltaPSI_50_vs_36, deltaPSI_36_vs_50)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_50_vs_36, aes(x=deltaPSI_50_vs_36, deltaPSI_36_vs_50), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_50_vs_36_sig, aes(x=deltaPSI_50_vs_36_sig, y=deltaPSI_36_vs_50_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_50_vs_36_sig, aes(x=deltaPSI_50_vs_36_sig, y=deltaPSI_36_vs_50_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("\u0394\u03A8 50 nt read length") +
  ylab("\u0394\u03A8 36 nt read length") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))
```

# 5 - Influence on the distribution of alternative splice events types (Fig 5)

Since there are alternative splicing events with different numbers of junctions, they are treated differently in terms of significance. 

Alternative splicing events with two junctions (A3SS, A5SS, AFE, ALE, IR) are considered significantly regulated if at least one junction has a dPSI of at least 0.05 and a probability changing value of at least 0.9. Events with 4 junctions (CE and MEE) are considered significantly regulated if either both junctions starting from C1 or both junctions starting from C2 fulfill the regulation criteria. 

```{r}
# load voila modulize files 
modDir = "Modulize/read_length"
modFiles = list.files(path = modDir, pattern = ".tsv")
setwd(modDir)

# list events with 2 Junctions separately for each read length 
events_2J = lapply(modFiles[1:20], loadFiles)
events_4J = lapply(modFiles[21:28], loadFiles)

# function to filter out significant events with 2 junctions
# significant events (dpsi >= 0.05, probability changing >= 0.9)
filter2JEvents <- function(eventTable) {
  eventTable = eventTable[abs(eventTable$colnames(eventTable)[21]) >= 0.05 & eventTable$colnames(eventTable)[22] >= 0.9, ] 
  eventTable %>% group_by(event_id) %>% slice(1)
}

# function to filter out significant events with 4 junctions
filter4JEvents <- function(eventTable) {
  eventTable %>% 
    group_by(event_id) %>%
    summarise(significant = (((abs(colnames(eventTable)[21][1]) >= 0.05 & colnames(eventTable)[22][1] >= 0.9) &
    (abs(colnames(eventTable)[21][2]) >= 0.05 & colnames(eventTable)[22][2] >= 0.9)) |
  ((abs(colnames(eventTable)[21][3]) >= 0.05 & colnames(eventTable)[22][3] >= 0.9) &
    (abs(colnames(eventTable)[21][4]) >= 0.05 & colnames(eventTable)[22][4] >= 0.9)))) %>%
    filter(significant == TRUE)
}

# filter out significant events
significantEvents_2J = lapply(events_2J, filter2JEvents)
significantEvents_4J = lapply(events_4J, filter4JEvents)
names(significantEvents_2J) = c("A3SS_100", "A3SS_75", "A3SS_50", "A3SS_36", 
                                      "A5SS_100", "A5SS_75", "A5SS_50", "A5SS_36", 
                                      "AFE_100", "AFE_75", "AFE_50", "AFE_36", 
                                      "ALE_100", "ALE_75", "ALE_50", "ALE_36", 
                                      "IR_100", "IR_75", "IR_50", "IR_36")
names(significantEvents_4J) = c("CE_100", "CE_75", "CE_50", "CE_36",
                                "MEE_100", "MEE_75", "MEE_50", "MEE_36")

# data frames with number of event types for each read length
numberOfEventTypes_100_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_100), 
                                             nrow(significantEvents_2J$A5SS_100),
                                             nrow(significantEvents_2J$AFE_100), 
                                             nrow(significantEvents_2J$ALE_100), 
                                             nrow(significantEvents_2J$IR_100), 
                                             nrow(significantEvents_4J$CE_100),
                                             nrow(significantEvents_4J$MEE_100)))

numberOfEventTypes_U2AF2_75_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_75), 
                                             nrow(significantEvents_2J$A5SS_75),
                                             nrow(significantEvents_2J$AFE_75), 
                                             nrow(significantEvents_2J$ALE_75), 
                                             nrow(significantEvents_2J$IR_75), 
                                             nrow(significantEvents_4J$CE_75),
                                             nrow(significantEvents_4J$MEE_75)))

numberOfEventTypes_50_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_50), 
                                             nrow(significantEvents_2J$A5SS_50),
                                             nrow(significantEvents_2J$AFE_50), 
                                             nrow(significantEvents_2J$ALE_50), 
                                             nrow(significantEvents_2J$IR_50), 
                                             nrow(significantEvents_4J$CE_50),
                                             nrow(significantEvents_4J$MEE_50)))

numberOfEventTypes_25_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_36), 
                                             nrow(significantEvents_2J$A5SS_36),
                                             nrow(significantEvents_2J$AFE_36), 
                                             nrow(significantEvents_2J$ALE_36), 
                                             nrow(significantEvents_2J$IR_36), 
                                             nrow(significantEvents_4J$CE_36),
                                             nrow(significantEvents_4J$MEE_36)))


# combine event type results of each read length into one data frame
numberOfEventTypes_100_df$readLength = "100"
numberOfEventTypes_U2AF2_75_df$readLength = "075"
numberOfEventTypes_50_df$readLength = "050"
numberOfEventTypes_36_df$readLength = "036"
numberOfEventTypes_df = rbind(numberOfEventTypes_100_df, 
                                    numberOfEventTypes_U2AF2_75_df, 
                                    numberOfEventTypes_50_df, 
                                    numberOfEventTypes_36_df)

# number of event types for each read length as bar charts
numberOfEventTypes_cols = 
  ggplot(numberOfEventTypes_df, aes(x = event, y = n, fill = readLength)) +
  geom_col(position = "dodge") +
  xlab("event type") +
  ylab("# events") +
  scale_fill_manual(values = c("036" = "plum2", "050" = "brown2", "075" = "cyan3", "100" = "midnightblue"), 
                    name = "read length [nt]", 
                    labels = c("36", "50", "75", "100")) +
  theme(plot.title = element_text(face = "bold")) 
 
# fraction of event types for each read length as stacked bar chart 
# amount of events log10 transformed
numberOfEventTypes_StackedBarChart = 
  ggplot(numberOfEventTypes_df, aes(x = readLength, y = n, fill = event)) +
  geom_col(position = "fill") +
  scale_y_log10() +
  xlab("read length [nt]") +
  ylab("# events (log10)") +
  scale_x_discrete(labels = c("36", "50", "75", "100")) +
  scale_fill_brewer(palette = "YlGnBu", name = "Event") +
  theme(plot.title = element_text(face = "bold"))
```

# Session info
```{r}
sessionInfo()
```

