---
title: "4 Metaanalysis of commonly used characteristics"
author: "Melina"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-overflow: scroll
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    
    toc: true
    toc-depth: 3
    toc-location: left
    toc-expand: false
    number-sections: true
    
    theme: sandstone
    fontsize: 11pt
    linestretch: 1.5
        
    cap-location: margin
    crossref:
      fig-title: Fig
    
    embed-resources: true
    link-external-newwindow: true
    smooth-scroll: true
    
    execute:
      echo: true
      warning: false
      cache: false

---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=FALSE, tidy.opts=list(width.cutoff=80))
```

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
library(knitr)

theme_set(theme_bw())
out <- "/Users/melinaklostermann/Documents/projects/xx_students/Annika_majiq_test/Ladwig_et_al_2024/4-out/"
```

- This report pulls metadata information from encode for all shRNA RNAseq, total RNAseq and polyA+ RNAseqs from encode.
- It then looks at the number of reads per sample and the mean number of reads per experiement.
- Finally, the GO terms splicing and RNA-binding are added for the shRNA targets to specifically look at the read numbers for splicing factors.

# Get metadata from encode

```{r eval = F}
# where to look
base_url <- "https://www.encodeproject.org/"

# ------------------
# shRNA RNAseq
# ------------------
## see https://www.encodeproject.org/help/rest-api/ for more info
query <- "search/?type=File&file_format=fastq&assay_title=shRNA%20RNA-seq&frame=object&limit=all"  
## this is a URI query string, %20 encodes an empthy space

# Make API request (API = Application Programming Interface)
response <- httr::GET(paste0(base_url, query), add_headers(accept = "application/json"))

# Parse JSON response (JSON = JavaScript Object Notation)
data <- jsonlite::fromJSON(content(response, as = "text", encoding = "UTF-8"))

# Most metadata is stored in data$`@graph`
#data$`@graph`$read_count

all_encode_shRNAseq_fastqs <- data$`@graph` %>% subset(status == "released")

# ------------------
# total RNAseq
# ------------------

query <- "search/?type=File&file_format=fastq&assay_title=total%20RNA-seq&frame=object&limit=all" 
## %20 encodes an empthy space

# Make API request
response <- GET(paste0(base_url, query), add_headers(accept = "application/json"))

# Parse JSON response
data <- fromJSON(content(response, "text", encoding = "UTF-8"))


all_encode_totalRNAseq_fastqs <- data$`@graph` %>% subset(status == "released")


# ------------------
# polyA RNAseq
# ------------------

query <- "search/?type=File&file_format=fastq&assay_title=polyA%20plus%20RNA-seq&frame=object&limit=all" 
## %20 encodes an empthy space

# Make API request
response <- GET(paste0(base_url, query), add_headers(accept = "application/json"))

# Parse JSON response
data <- fromJSON(content(response, "text", encoding = "UTF-8"))


all_encode_polyARNAseq_fastqs <- data$`@graph` %>% subset(status == "released")

saveRDS(all_encode_shRNAseq_fastqs, paste0(out, "meta_shRNAseq.rds"))
saveRDS(all_encode_totalRNAseq_fastqs, paste0(out, "meta_totalRNAseq.rds"))
saveRDS(all_encode_polyARNAseq_fastqs, paste0(out, "meta_polyARNAseq.rds"))

```



```{r}
all_encode_shRNAseq_fastqs <- readRDS(paste0(out, "meta_shRNAseq.rds"))
all_encode_totalRNAseq_fastqs <- readRDS(paste0(out, "meta_totalRNAseq.rds"))
all_encode_polyARNAseq_fastqs <- readRDS(paste0(out, "meta_polyARNAseq.rds"))

```

## Read numbers of each sample for shRNA, total and poly+ RNAseq separate

```{r}

# ------------------
# plot read counts
# ------------------

ggplot(all_encode_shRNAseq_fastqs, aes(x = read_count)) +
  geom_histogram()+
  scale_x_log10()+
  ggtitle("shRNA RNAseq")


ggplot(all_encode_totalRNAseq_fastqs, aes(x = read_count)) +
  geom_histogram()+
  scale_x_log10()+
  ggtitle("total RNAseq")

ggplot(all_encode_polyARNAseq_fastqs, aes(x = read_count)) +
  geom_histogram()+
  scale_x_log10()+
  ggtitle("polyA plus RNAseq")


```

# Adding more information
## Adding information on the year 

The year is stored inside the date created column and needs to be extracted.

```{r}
# -------------------
# add year info 
# -------------------

all_encode_shRNAseq_fastqs <- all_encode_shRNAseq_fastqs %>% 
  rowwise() %>%
  mutate( year = as.numeric(unlist(strsplit(date_created, split = "-"))[[1]]))

all_encode_totalRNAseq_fastqs <- all_encode_totalRNAseq_fastqs %>% 
  rowwise() %>%
  mutate( year = as.numeric(unlist(strsplit(date_created, split = "-"))[[1]]))

all_encode_polyARNAseq_fastqs <- all_encode_polyARNAseq_fastqs %>% 
  rowwise() %>%
  mutate( year = as.numeric(unlist(strsplit(date_created, split = "-"))[[1]]))


table(all_encode_shRNAseq_fastqs$year)

table(all_encode_totalRNAseq_fastqs$year)

table(all_encode_polyARNAseq_fastqs$year)

```

## Adding information on the organism

The organism is not given in the metadata of type "file" that was obtained above, but in the metadata of type "experiment". This is pulled here and both are combined by the accession number.

```{r eval=F}
# -------------------
# add organism
# -------------------

# shRNAseq
query <- "search/?type=Experiment&assay_title=shRNA%20RNA-seq&limit=all"
response <- GET(paste0(base_url, query), add_headers(accept = "application/json"))
exp_data <- fromJSON(content(response, "text", encoding = "UTF-8"))


exp_data_small <- exp_data$`@graph` %>% dplyr::select(c("accession", "biosample_summary")) %>%
  rowwise() %>%
  mutate(organism  = paste(unlist(strsplit(biosample_summary, split = " "))[[1]], unlist(strsplit(biosample_summary, split = " "))[[2]])) 
                                                           

all_encode_shRNAseq_fastqs <- all_encode_shRNAseq_fastqs %>%
  rowwise() %>%
  mutate(exp_id  = unlist(strsplit(dataset, split = "/"))[[3]]) %>%
  left_join(exp_data_small, by = c(exp_id = "accession"))

# totalRNAseq
query <- "search/?type=Experiment&assay_title=total%20RNA-seq&limit=all"
response <- GET(paste0(base_url, query), add_headers(accept = "application/json"))
exp_data <- fromJSON(content(response, "text", encoding = "UTF-8"))


exp_data_small <- exp_data$`@graph` %>% dplyr::select(c("accession", "biosample_summary")) %>%
  rowwise() %>%
  mutate(organism  = paste(unlist(strsplit(biosample_summary, split = " "))[[1]], unlist(strsplit(biosample_summary, split = " "))[[2]])) 
                                                           

all_encode_totalRNAseq_fastqs <- all_encode_totalRNAseq_fastqs %>%
  rowwise() %>%
  mutate(exp_id  = unlist(strsplit(dataset, split = "/"))[[3]]) %>%
  left_join(exp_data_small, by = c(exp_id = "accession"))


# polARNAseq
query <- "search/?type=Experiment&assay_title=polyA%20plus%20RNA-seq&limit=all"
response <- GET(paste0(base_url, query), add_headers(accept = "application/json"))
exp_data <- fromJSON(content(response, "text", encoding = "UTF-8"))


exp_data_small <- exp_data$`@graph` %>% dplyr::select(c("accession", "biosample_summary")) %>%
  rowwise() %>%
  mutate(organism  = paste(unlist(strsplit(biosample_summary, split = " "))[[1]], unlist(strsplit(biosample_summary, split = " "))[[2]])) 
                                                           

all_encode_polyARNAseq_fastqs <- all_encode_polyARNAseq_fastqs %>%
  rowwise() %>%
  mutate(exp_id  = unlist(strsplit(dataset, split = "/"))[[3]]) %>%
  left_join(exp_data_small, by = c(exp_id = "accession"))


saveRDS(all_encode_shRNAseq_fastqs, paste0(out, "meta_shRNAseq_withExp.rds"))
saveRDS(all_encode_totalRNAseq_fastqs, paste0(out, "meta_totalRNAseq_withExp.rds"))
saveRDS(all_encode_polyARNAseq_fastqs, paste0(out, "meta_polyARNAseq_withExp.rds"))

```


```{r}

all_encode_shRNAseq_fastqs <- readRDS(paste0(out, "meta_shRNAseq_withExp.rds"))
all_encode_totalRNAseq_fastqs <- readRDS(paste0(out, "meta_totalRNAseq_withExp.rds"))
all_encode_polyARNAseq_fastqs <- readRDS(paste0(out, "meta_polyARNAseq_withExp.rds"))

# look at numbers
table(all_encode_shRNAseq_fastqs$organism)
table(all_encode_totalRNAseq_fastqs$organism)
table(all_encode_polyARNAseq_fastqs$organism)



```

## Combination of shRNA, total and polyA+ RNAseq

```{r}
# -------------------
# combine
# -------------------
all_encode_RNAseq <- rbind(all_encode_shRNAseq_fastqs %>% 
                             dplyr::select(c("exp_id", "organism", "read_count", "year", "assay_title", "biological_replicates_formatted", "biosample_summary", "target" )),
                           
                           all_encode_totalRNAseq_fastqs %>% 
                             dplyr::select(c("exp_id", "organism", "read_count", "year", "assay_title", "biological_replicates_formatted", "biosample_summary" )) %>% 
                                      mutate(target = ""),
                           
                           all_encode_polyARNAseq_fastqs %>% 
                             dplyr::select(c("exp_id", "organism", "read_count", "year", "assay_title", "biological_replicates_formatted", "biosample_summary" )) %>% 
                                      mutate(target = ""))




ggplot(all_encode_RNAseq, aes(x = read_count)) +
  geom_histogram()+
  scale_x_log10()+
  ggtitle("all RNAseq")


ggplot(all_encode_RNAseq, aes(x = as.numeric(read_count), fill = year)) +
  geom_histogram()+
  scale_x_log10()+
  ggtitle("all RNAseq")

ggplot(all_encode_RNAseq, aes(y = as.numeric(read_count), x = as.character(year), group =year)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  ggtitle("all RNAseq")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

ggplot(all_encode_RNAseq, aes(y = as.numeric(read_count), x = as.character(year), group =year)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  facet_wrap(~organism)+
  ggtitle("all RNAseq")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

ggplot(all_encode_RNAseq, aes(y = as.numeric(read_count), x = as.character(year), group =year)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  facet_wrap(~assay_title)+
  ggtitle("all RNAseq")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

ggplot(all_encode_RNAseq, aes(y = as.numeric(read_count), x = "all")) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  ggtitle("all RNAseq")+
  xlab("")

```


## Same plots for mean read number per experiment

```{r}
# --------------------
# Mean read number per experiment
# ---------------------

all_encode_RNAseq <- group_by(all_encode_RNAseq, exp_id) %>%
  mutate(mean_read_count = mean(read_count))


ggplot(all_encode_RNAseq, aes(y = as.numeric(mean_read_count), x = as.character(year), group =year)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

ggplot(all_encode_RNAseq, aes(y = as.numeric(mean_read_count), x = as.character(year), group =year)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  facet_wrap(~organism)+
  ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

ggplot(all_encode_RNAseq, aes(y = as.numeric(mean_read_count), x = as.character(year), group =year)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "darkorange", size = 1)+
  scale_y_log10()+
  facet_wrap(~assay_title)+
  ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

```

```{r fig.width= 16, fig.height = 4}
# Plot for paper
all_encode_RNAseq_int_org1 <- subset(all_encode_RNAseq, organism %in% c("Homo sapiens", "Mus musculus"))
all_encode_RNAseq_int_org1$organism <- factor(all_encode_RNAseq_int_org1$organism, levels = c("Homo sapiens", "Mus musculus"))

all_encode_RNAseq_int_org2 <- subset(all_encode_RNAseq, organism %in% c( "Caenorhabditis elegans", "Drosophila melanogaster"))
all_encode_RNAseq_int_org2$organism <- factor(all_encode_RNAseq_int_org2$organism, levels = c("Drosophila melanogaster", "Caenorhabditis elegans"))

p0 = ggplot(all_encode_RNAseq_int_org1, aes(y = as.numeric(mean_read_count), x = as.character(year), group =year)) +
  geom_hline(yintercept = 30000000,  size = 1.5)+
  geom_hline(yintercept = 30000000, color = "#064644", size = 1)+
  geom_hline(yintercept = 1000000,  size = 1.5)+
  geom_hline(yintercept = 1000000, color = "#C7F8F5", size = 1)+
  geom_boxplot( fill = "white")+
  scale_y_log10()+
  facet_wrap(~organism, ncol = 4, scales = "free_x")+
  #ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("Year")+
  ylab("Read count \n mean of replicates per experiment [log10]")

p0

ggsave(p0, filename = paste0(out, "Encode_read_count_hs_mm.pdf"), width = 10, height = 3)


p01 = ggplot(all_encode_RNAseq_int_org2, aes(y = as.numeric(mean_read_count), x = as.character(year), group =year)) +
  ggrastr::rasterise( gghalves::geom_half_point(alpha = 0.6, color = "black", size = 0.8), dpi = 300)+
  gghalves::geom_half_boxplot(outlier.shape = NA, fill = "slategrey", alpha = 0.5)+
  geom_hline(yintercept = 30000000, color = "turquoise3", size = 1)+
  geom_hline(yintercept = 1000000, color = "aquamarine3", size = 1)+
  scale_y_log10()+
  facet_wrap(~organism, ncol = 4, scales = "free_x")+
  #ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("Year")+
  ylab("Read count \n mean of replicates per experiment [log10]")

p01

ggsave(p01, filename = paste0(out, "Encode_read_count_dm_ce.pdf"), width = 5, height = 3)

# p01 = ggplot(all_encode_RNAseq_int_org, aes(y = as.numeric(mean_read_count), x = as.character(year), group =year)) +
#   geom_hline(yintercept = 30000000,  size = 1.5)+
#   geom_hline(yintercept = 30000000, color = "#064644", size = 1)+
#   geom_hline(yintercept = 1000000,  size = 1.5)+
#   geom_hline(yintercept = 1000000, color = "#C7F8F5", size = 1)+
#   geom_boxplot( fill = "white")+
#   scale_y_log10()+
#   facet_wrap(~organism, ncol = 4, scales = "free_x")+
#   #ggtitle("all RNAseq", subtitle = "mean per replicate")+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
#   xlab("Year")+
#   ylab("Read count \n mean of replicates per experiment [log10]")
# 
# p01
# 
# ggsave(p01, filename = paste0(out, "Encode_read_count_altcol.pdf"), width = 10, height = 3)
# 
# group_by(all_encode_RNAseq_int_org, organism) %>%
# summarise( median(as.numeric(mean_read_count), na.rm  = T))

```


# What are good data sets?

```{r}

all_encode_RNAseq %>% subset(!is.na(read_count) ) %>% arrange(desc(mean_read_count)) %>% head(10) %>% kable()

all_encode_RNAseq %>% subset(!is.na(read_count) & assay_title == "shRNA RNA-seq" ) %>% arrange(desc(mean_read_count)) %>% head(10) %>% kable()


```


# Mark splicing factors & RBPs




```{r}
# -----------------
# get GO lists
# -----------------
library(BiocSet)
library(GO.db)
library(org.Hs.eg.db)

# Splicing
#---------------

# Get all BP GO terms
go <- go_sets(org.Hs.eg.db, "SYMBOL", go = c("GO", "GOID"), ontology = "BP")

map <- map_add_set(go, GO.db, "GOID", "TERM")
go <-  go %>% mutate_set(term = map)
s <- es_set(go)

# find term for nucleus and nucleolus

# s %>% subset(grepl(s$term, pattern = "splicing"))
s %>% subset(grepl(s$term, pattern = "binding"))

splicing_set_id <- "GO:0008380"


# get sets
es <- es_elementset(go)

splicing_set <- es %>% subset(set == splicing_set_id)

# RNA binding
# --------------

go <- go_sets(org.Hs.eg.db, "SYMBOL", go = c("GO", "GOID"), ontology = "MF")

map <- map_add_set(go, GO.db, "GOID", "TERM")
go <-  go %>% mutate_set(term = map)
s <- es_set(go)

# s %>% subset(grepl(s$term, pattern = "RNA binding"))

rnaBinding_set_id <- "GO:0003723"

# get sets
es <- es_elementset(go)

rnaBinding_set <- es %>% subset(set == rnaBinding_set_id)


# ---------------------
# add to metadata
# --------------------

# only applicable for shRNA RNAseq
shRNAseq <- all_encode_RNAseq %>% subset(!is.na(read_count) & (organism == "Homo sapiens")) %>%
  mutate(target = unlist(strsplit(target, split = "/"))[3],
         target = unlist(strsplit(target, split = "-"))[1])


shRNAseq$splicing <- shRNAseq$target %in% splicing_set$element
table(shRNAseq$splicing)

shRNAseq$rnaBinding <- shRNAseq$target %in% rnaBinding_set$element
table(shRNAseq$rnaBinding)


ggplot(shRNAseq, aes(y = as.numeric(mean_read_count), x = splicing, colour = splicing)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "orange", size = 1)+
  scale_y_log10()+
  scale_color_manual(values = c("grey30", "dodgerblue3"))+
  ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("Year")+
  ylab("Read count \n mean of replicates per experiment [log10]")

ggplot(shRNAseq, aes(y = as.numeric(mean_read_count), x = rnaBinding, colour = rnaBinding)) +
  gghalves::geom_half_point(alpha = 0.2)+
  gghalves::geom_half_boxplot()+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "orange", size = 1)+
  scale_y_log10()+
  scale_color_manual(values = c("grey30", "dodgerblue3"))+
  ggtitle("all RNAseq", subtitle = "mean per replicate")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5))+
  xlab("")

```

```{r}
# plots for paper

p = ggplot(shRNAseq, aes(y = as.numeric(mean_read_count), x = splicing, colour = splicing)) +
  gghalves::geom_half_boxplot()+
  ggrastr::rasterise( gghalves::geom_half_point(alpha = 0.2,size = 0.8), dpi = 300)+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "orange", size = 1)+
  scale_y_log10()+
  scale_color_manual(values = c("grey50", "dodgerblue3"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "none")+
  xlab("")+
  ylab("Read count \n mean of replicates per experiment [log10]")

ggsave(p, filename = paste0(out, "Encode_read_count_splicing.pdf"), width = 6, height = 6, units = "cm")

p1 = ggplot(shRNAseq, aes(y = as.numeric(mean_read_count), x = rnaBinding, colour = rnaBinding)) +
  gghalves::geom_half_boxplot()+
    ggrastr::rasterise( gghalves::geom_half_point(alpha = 0.2, size = 0.8), dpi = 300)+
  geom_hline(yintercept = 30000000, color = "darkred", size = 1)+
  geom_hline(yintercept = 1000000, color = "orange", size = 1)+
  scale_y_log10()+
  scale_color_manual(values = c("grey50", "dodgerblue3"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "none")+
  xlab("")+
  ylab("Read count \n mean of replicates per experiment [log10]")

ggsave(p1, filename = paste0(out, "Encode_read_count_rnaBinding.pdf"), width = 6, height = 6, units = "cm")

```

## Best splicing data sets

```{r}
shRNAseq %>% subset(splicing) %>% arrange(desc(mean_read_count)) %>% head(20) %>% kable()

```







# Session Info

```{r}

sessionInfo()

```