---
title: "5 Metaanalysis of commonly used characteristics in SRA"
author: "Melina"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-overflow: scroll
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    
    toc: true
    toc-depth: 3
    toc-location: left
    toc-expand: false
    number-sections: true
    
    theme: sandstone
    fontsize: 11pt
    linestretch: 1.5
        
    cap-location: margin
    crossref:
      fig-title: Fig
    
    embed-resources: true
    link-external-newwindow: true
    smooth-scroll: true
    
    execute:
      echo: true
      warning: false
      cache: false

---

```{r}
library(httr)
library(jsonlite)
library(tidyverse)
library(knitr)
library(rentrez)

set_entrez_key("1ac0cb6caefae06168139145ce4a97f0c509") # allows to access more data

theme_set(theme_bw())
out <- "/Users/melinaklostermann/Documents/projects/xx_students/Annika_majiq_test/Ladwig_et_al_2024/5-out/"
```

```{r eval=FALSE, include=F}

library(GEOmetadb)
geometadbfile <- "/Users/melinaklostermann/Documents/projects/PublicData/GEO/GEOmetadb_2024_05_01.sqlite"
# SQL = Structured Query language
con <- dbConnect(SQLite(), geometadbfile)

dbListFields(con,'gsm')                                                                                                         

rs <- dbGetQuery(con,paste("select ID,title,description,series_id,gsm,organism_ch1,characteristics_ch1,treatment_protocol_ch1,source_name_ch1,contact,submission_date,supplementary_file,data_row_count from gsm where", 
                           "title like '%RNA-seq%'" ,sep=" "))

# does not contain read number

```

# Checking SRA for read counts of RNAseqs

- Works over the NCBI browser via the rentrez package.
- I can look at mak 300 samples at once.

```{r eval=FALSE}

entrez_dbs()
entrez_db_summary("sra")

entrez_db_searchable("sra")

s <- entrez_search(db = "sra", term = "(RNA-seq [STRA]) AND Homo Sapiens [ORGN] AND 2024 [PDAT] AND TRANSCRIPTOMIC [SRC]", retmax = 300)
entrez_db_links("sra")

f <- entrez_fetch(db="sra", id = s$ids, retmode = "xml", rettype="runinfo", parsed=TRUE, use_history=TRUE)

meta_data <- lapply(f,function(x) XML::xmlToDataFrame(x)) %>% as.data.frame()

saveRDS(meta_data, paste0(out, "sra_RNAseq_300_", Sys.Date(), ".rds"))
```

```{r}
meta_data <- readRDS(paste0(out, "sra_RNAseq_300_2025-01-17.rds"))

meta_data <- meta_data %>% subset(spots != 0)

ggplot(meta_data, aes(x = 1, y = as.numeric(spots)))+
  gghalves::geom_half_boxplot()+
  gghalves::geom_half_point()+
  scale_y_log10()+
  geom_hline(yintercept = 30000000, color = "darkorange")+
  ggtitle("300 RNAseq data sets from 2024")
```

# Specific splicing data sets

```{r}
# Only containing the term splicing
s <- entrez_search(db = "sra", term = "(RNA-seq [STRA]) AND (Homo Sapiens [ORGN] AND 2024 [PDAT] AND splicing AND TRANSCRIPTOMIC [SRC])", retmax = 300)
entrez_db_links("sra")

f <- entrez_fetch(db="sra", id = s$ids, retmode = "xml", rettype="runinfo", parsed=TRUE, use_history=TRUE)

meta_data <- lapply(f,function(x) XML::xmlToDataFrame(x)) %>% as.data.frame()

saveRDS(meta_data, paste0(out, "sra_RNAseq_splicing_300_", Sys.Date(), ".rds"))
```

```{r}
meta_data <- readRDS(paste0(out, "sra_RNAseq_splicing_300_2025-01-17.rds"))
meta_data <- meta_data %>% subset(spots != 0)

ggplot(meta_data, aes(x = 1, y = as.numeric(spots)))+
  gghalves::geom_half_boxplot()+
  gghalves::geom_half_point()+
  scale_y_log10()+
  geom_hline(yintercept = 30000000, color = "darkorange")+
  ggtitle("300 RNAseq data sets from 2024 connected to splicing")


```



# Ideas for good data sets for an additional analysis

```{r}
data_study_wise <- meta_data %>% subset(avgLength >= 300) %>% 
  group_by(SRAStudy) %>%
  summarise(SRAStudy = SRAStudy,
            n_samples = n(),
            min_spots = min(spots),
            ) %>%
  arrange(min_spots)

data_study_wise

```

Two data sets with > 100 Mio reads:

SRP534967: THUMPD3 regulates alternative splicing of ECM transcripts in human lung cancer cells and promotes proliferation and migration (human)
SRP523909: Characterization of aberrant splicing reveals CLK1 as a candidate oncogenic dependency in pediatric high-grade gliomas (human) --> Morpholino treatment


