---
title: "Impact of replicate number on alternative splicing analysis (endodermall cell vs H1 cell line)"
author: "Annika Ladwig"
output:
  prettydoc::html_pretty:
    theme: architect
---

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE,fig.keep='all'}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE,fig.keep='all'}
# LIBRARIES 
library(dplyr)
library(ggplot2)
library(stringr)
library(ggpubr)
library("RColorBrewer")
library("ggVennDiagram")
library("wesanderson")
library(ggrastr)
library("colorspace")
#library(ggecdf)
library(UpSetR)
library(halfmoon)

# MY PARMETERS
probabilityThreshold = 0.9

# MY THEME
source("C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/theme_paper.R")
theme_set(theme_paper())

# MY FUNCTIONS
# function to load files as data frames
loadFiles <- function(x){
  x = read.table(file = x,
                 header = TRUE,
                 sep = "\t",
                 stringsAsFactors = FALSE)
}

# function to count semicolons of column "mean_dpsi_per_lsv_junction"
# to find out the number of junctions per LSV
# store this information in new column "nJunctions"
countJunctions <- function(x) {
    junctionsPerLSV = x$mean_dpsi_per_lsv_junction %>%
   str_count(pattern = ";") + 1
  x$nJunctions = junctionsPerLSV
  x
}

# count the amount of LSVs with 2, 3, 4, ... junctions
# create data frame with this information
complexityLSVs <- function(x) {
   junctionsPerLSV = x$nJunctions %>%
    table(dnn="junctionsPerLSV") %>%
    as.data.frame
   junctionsPerLSV
}

# function to extract max dpsi for each LSV and add new column with this information
extractMaxDPSIs <- function(x) {
  maxDeltaPSIs = x$mean_dpsi_per_lsv_junction %>%
    str_split(pattern = ";") %>%
    sapply(FUN = function(dPSIs){
      return(
        dPSIs %>% as.numeric %>% abs %>% max
      )
    })
  x$maxDeltaPSI = maxDeltaPSIs
  x
}

# function extract max prob. for each LSV and add a new column with this information
extractMaxProbs <- function(x) {
  maxProbs = x$probability_changing %>%
    str_split(pattern=";") %>%
    sapply(FUN=function(probs){
      return(
        probs %>% as.numeric %>% max
      )
    })
  x$maxProb = maxProbs
  x
}
```

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE,fig.keep='all'}
# LOAD FILE SETS (read length 100 nt)

# 2 samples
# load voila files as list
two_samples_Dir = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/endodermal_cell_vs_H1_cell_line/MAJIQ/2_samples/"

two_samples_files = list.files(path = two_samples_Dir, pattern = ".tsv")
setwd(two_samples_Dir)

# load dpsi tsv files as data frames
# & name list elements
two_samples_dfs = lapply(two_samples_files, loadFiles)
names(two_samples_dfs) = c("sample_1_2", "sample_1_3", "sample_2_3")


# 3 samples
# load voila files as list
three_samples_Dir = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/endodermal_cell_vs_H1_cell_line/MAJIQ/3_samples"

three_samples_files = list.files(path = three_samples_Dir, pattern = ".tsv")
setwd(three_samples_Dir)

# load dpsi tsv files as data frames
# & name list elements
three_samples_dfs = lapply(three_samples_files, loadFiles)
names(three_samples_dfs) = c("sample_1_2_3")
```


# 2.1 - Influence on LSV detection (Fig 2A)

The influence of replicate number on LSV detection is determined by counting the LSV hits (rows) in the TSV files received from MAJIQ for each replicate number. 

```{r}
# put all sample variations in one list
sample_sizes_df = append(two_samples_dfs, list(three_samples_dfs$sample_1_2_3))

two_samples_dfs[1]

# detected LSVs
nLSVs_sample_1_2 = nrow(sample_sizes_df[[1]])
nLSVs_sample_1_3 = nrow(sample_sizes_df[[2]])
nLSVs_sample_2_3 = nrow(sample_sizes_df[[3]])
nLSVs_sample_1_2_3 = nrow(sample_sizes_df[[4]])

cat("two samples", fill = TRUE)
nLSVs_sample_1_2
nLSVs_sample_1_3
nLSVs_sample_2_3

cat("three samples", fill = TRUE)
nLSVs_sample_1_2_3

# create dataframe with nLSVs 
nLSVs_df = data.frame(nSamples=c("2", "2", "2", "3"),
                      nLSVs=c(nLSVs_sample_1_2, nLSVs_sample_1_3, nLSVs_sample_2_3, nLSVs_sample_1_2_3))

# number of LSVs bar chart
nLSVs_sampleSize = ggplot(nLSVs_df, aes(nSamples, y = nLSVs))+
  stat_summary(fun = "mean",
               geom = "bar",
               size = 0.5,
               show.legend = FALSE,
               mapping = aes(fill=nSamples)) +
  scale_fill_manual(name = "", values = c("#EFC4E4", "#D065B5")) +
  ylim(0, 50500) +
  geom_point(alpha = 0.6, color = darken("#EFC4E4", 0.25)) +
  xlab("# replicates") +
  ylab("# nLSVs") +
  ggtitle("D") +
  scale_x_discrete(labels = c("75", "100")) +
  geom_text(aes(label=nLSVs), position=position_dodge(width=0.9), vjust=1)

nLSVs_sampleSize

# save ggplots
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/nLSVs_sampleSize.pdf", plot = nLSVs_sampleSize, width = 2.46, height = 6, units = "cm")
```


# 2.2 - Influence on LSV complexity (Fig 2B)

The LSV complexity describes how many junctions an LSV has. To measure the influence on this, the LSVs with 2, 3, 4 or more than 4 junctions are counted for each replicate number. 

```{r}
# create new list with the data frames that contain the information about the complexity of the LSVs
sample_sizes_df = lapply(sample_sizes_df, countJunctions)
sample_sizes_complexityLSVs = lapply(sample_sizes_df, complexityLSVs)

# add new column for number of samples to differ the data later
# when merging single data frames to one
sample_sizes_complexityLSVs[[4]]$n_samples = "3"

# build mean data frame out of data frames for 2 samples 
junctionsPerLSV = sample_sizes_complexityLSVs[[1]]$junctionsPerLSV
Freq = c()
for (i in 1:22){
  Freq = append(Freq, mean(c(sample_sizes_complexityLSVs[[1]]$Freq[i], sample_sizes_complexityLSVs[[2]]$Freq[i], sample_sizes_complexityLSVs[[3]]$Freq[i])))
}
two_samples_complexityLSVs = data.frame(junctionsPerLSV = junctionsPerLSV, 
                                        Freq = Freq,
                                        n_samples = 2)

# create data frame out of all sample sizes to build a stacked bar chart
nJunctions_df = rbind(two_samples_complexityLSVs, sample_sizes_complexityLSVs[[4]]) 

# all LSVs with junctions > 4 are grouped into one category
# so that all LSVs with > 4 junctions get the same color in following stacked bar chart
for (i in 1:nrow(nJunctions_df)){
  if (as.integer(nJunctions_df[i, "junctionsPerLSV"]) >= 5) {
    nJunctions_df[i, "junctionsPerLSV"] = factor(5)
  }
}

# stacked bar chart to show complexity of LSVs for each sample size
# source: http://jessicacorman.weebly.com/news--musings/how-to-make-a-stacked-bar-chart-with-color-shading
StackedBarChart_nJunctions = ggplot(nJunctions_df, aes(x = n_samples, y = Freq)) +
  geom_bar(stat = "identity", aes(fill = n_samples), position = "fill", show.legend = FALSE) +
  scale_fill_manual(values = c("#EFC4E4", "#C76EB1"),
                    labels = c("2", "3"),
                    name = "# samples") + 
  geom_bar(stat = "identity", fill = "black", aes(alpha = junctionsPerLSV), position = "fill", show.legend = FALSE) +
  scale_alpha_manual(values = c(0, 0.3, 0.55, 0.8),
                     name = "junctions per LSV", 
                     breaks = c("2", "3", "4", "5"),
                     labels = c("2", "3", "4", "> 4")) +
  scale_x_discrete(labels = c("2", "3")) +
  xlab("#samples") +
  ylab("fraction LSVs [%]") +
  ggtitle("B") 

StackedBarChart_nJunctions

# save ggplots
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/complexityLSVs_sampleSize.pdf", plot = StackedBarChart_nJunctions, width = 2.26, height = 6, units = "cm")
```


# 3 - Influence on the significance of LSVs (Fig 3)

Influence of replicate number on the significance of LSVs is determined by plotting the fraction of significantly regulated LSVs detected for each replicate number. LSVs are considered significant if dPSI is at least 0.05 and probability changing is at least 0.9. 

For each LSV, the maximum dPSI and max probability changing is chosen. These usually belong to the same junction.

## Number and fraction of detected significant LSVs (Fig 3F)

```{r}
# extract max dpsi for each LSV and add new column with this information
sample_sizes_df = lapply(sample_sizes_df, extractMaxDPSIs)

# extract max prob. for each LSV and add a new column with this information
sample_sizes_df = lapply(sample_sizes_df, extractMaxProbs)

# add new column called "significant" for each sample combination which could be TRUE/FALSE 
# TRUE if maxProb >= 0.9 and maxDPSI >= 0.05
sample_sizes_df[[1]]$significant = sample_sizes_df[[1]]$maxProb >= probabilityThreshold & sample_sizes_df[[1]]$maxDeltaPSI >= 0.05
sample_sizes_df[[2]]$significant = sample_sizes_df[[2]]$maxProb >= probabilityThreshold & sample_sizes_df[[2]]$maxDeltaPSI >= 0.05
sample_sizes_df[[3]]$significant = sample_sizes_df[[3]]$maxProb >= probabilityThreshold & sample_sizes_df[[3]]$maxDeltaPSI >= 0.05
sample_sizes_df[[4]]$significant = sample_sizes_df[[4]]$maxProb >= probabilityThreshold & sample_sizes_df[[4]]$maxDeltaPSI >= 0.05


# output which LSVs are significant and which are not for sample size combination
table(sample_sizes_df[[1]]$significant)
table(sample_sizes_df[[2]]$significant)
table(sample_sizes_df[[3]]$significant)
table(sample_sizes_df[[4]]$significant)

# count significant and not significant LSVs for each sample combination
nSignificant_1_2 = sum(sample_sizes_df[[1]]$significant)
nNotSignificant_1_2 = (nLSVs_sample_1_2-sum(sample_sizes_df[[1]]$significant))
nSignificant_1_3 = sum(sample_sizes_df[[2]]$significant)
nNotSignificant_1_3 = (nLSVs_sample_1_3-sum(sample_sizes_df[[2]]$significant))
nSignificant_2_3 = sum(sample_sizes_df[[3]]$significant)
nNotSignificant_2_3 = (nLSVs_sample_2_3-sum(sample_sizes_df[[3]]$significant))
nSignificant_1_2_3 = sum(sample_sizes_df[[4]]$significant)
nNotSignificant_1_2_3 = (nLSVs_sample_1_2_3-sum(sample_sizes_df[[4]]$significant))

# calculate mean of regulated LSVs for sample size of 2
mean_regulated_2_samples = round(mean(c(nSignificant_1_2, nSignificant_1_3, nSignificant_2_3)))
mean_nLSVs_2_samples = round(mean(nLSVs_sample_1_2, nLSVs_sample_1_3, nLSVs_sample_2_3))
mean_notRegulated_2_samples =mean_nLSVs_2_samples - mean_regulated_2_samples

# create data frame with a column for the replicate number
# and a column indicating whether the associated LSV is significant or not 
nSignificantLSVs_df_relative = data.frame(comparison=c(rep("012", nrow(sample_sizes_df[[1]])),
                                                   rep("013", nrow(sample_sizes_df[[2]])), 
                                                   rep("023", nrow(sample_sizes_df[[3]])), 
                                                   rep("123", nrow(sample_sizes_df[[4]]))),
                                      significant=c(sample_sizes_df[[1]]$significant,
                                                    sample_sizes_df[[2]]$significant, 
                                                    sample_sizes_df[[3]]$significant, 
                                                    sample_sizes_df[[4]]$significant))

nSignificantLSVs_df_relative = data.frame(comparison=rep(c("2", "3"), each = 2),
                                          significant=c("TRUE 2", "FALSE", "TRUE 3", "FALSE"),
                                          n=c(mean_regulated_2_samples, mean_nLSVs_2_samples, 
                                              sum(sample_sizes_df[[4]]$significant), nrow(sample_sizes_df[[4]])))
nSignificantLSVs_df_relative

specie <- c(rep("sorgho" , 3) , rep("poacee" , 3) , rep("banana" , 3) , rep("triticum" , 3) )
condition <- rep(c("normal" , "stress" , "Nitrogen") , 4)
value <- abs(rnorm(12 , 0 , 15))
data <- data.frame(specie,condition,value)
data

# percentage stacked bar chart
# stacked bar chart showing fraction of significant LSVs among all LSVs for each replicate number  
nSignificantLSVs_bar_relative = ggplot(nSignificantLSVs_df_relative, aes(x = comparison, y = n, fill = significant)) +
  geom_bar(position = "fill", stat = "identity", show.legend = FALSE) +
  xlab("# samples") +
  ylab("fraction LSVs [%]") +
  coord_cartesian(ylim = c(0, 0.02)) +
  scale_fill_manual(name= "significant", values = c("lightgrey", "#EFC4E4", "#D065B5"), labels = c("FALSE", "TRUE")) +
  ggtitle("A") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))

nSignificantLSVs_bar_relative


# bar chart with absolute number of significant LSVs
# create dataframe with significant nLSVs 
nSignificantLSVs_df_absolute = data.frame(nSamples=c("2", "2", "2", "3"),
                      nSignificantLSVs=c(nSignificant_1_2, nSignificant_1_3, nSignificant_2_3, nSignificant_1_2_3))

nSignificantLSVs_bar_absolute = ggplot(nSignificantLSVs_df_absolute, aes(nSamples, nSignificantLSVs))+
  stat_summary(fun = "mean",
               geom = "bar",
               size = 0.5,
               show.legend = FALSE,
               mapping = aes(fill=nSamples)) +
  scale_fill_manual(name = "", values = c("#EFC4E4", "#D065B5")) +
  geom_point(alpha = 0.6, color = darken("#EFC4E4", 0.25)) +
  xlab("# samples") +
  ylab("# significant LSVs") +
  ggtitle("C") +
  geom_text(aes(label=nSignificantLSVs), position=position_dodge(width=0.9), vjust=1)
  
nSignificantLSVs_bar_absolute


# save plots
ggsave(filename = "nSignficantLSVs_sample_size_absolute.pdf", plot = nSignificantLSVs_bar_absolute, width = 2.2, height = 6, units = "cm", path = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots")
ggsave(filename = "nSignficantLSVs_sample_size_relative.pdf", plot = nSignificantLSVs_bar_relative, width = 2.41, height = 6, units = "cm", path = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots")
```

## Shared significant regulated LSVs between the different sample sizes (Fig 3G)

```{r}
# check number of regulated LSVs
# TRUE = 1, FALSE = 0
sum(sample_sizes_df[[1]]$significant)
sum(sample_sizes_df[[2]]$significant)
sum(sample_sizes_df[[3]]$significant)
sum(sample_sizes_df[[4]]$significant)

# extract all significant LSVs for each replicate number
significantLSVs_samples_12 = sample_sizes_df[[1]]$lsv_id[sample_sizes_df[[1]]$significant == TRUE]

significantLSVs_samples_13 = sample_sizes_df[[2]]$lsv_id[sample_sizes_df[[2]]$significant == TRUE]

significantLSVs_samples_23 = sample_sizes_df[[3]]$lsv_id[sample_sizes_df[[3]]$significant == TRUE]

significantLSVs_samples_123 = sample_sizes_df[[4]]$lsv_id[sample_sizes_df[[4]]$significant == TRUE]

significantLSVs_all = list("012" = c(significantLSVs_samples_12), "013" = c(significantLSVs_samples_13), "023" = c(significantLSVs_samples_23), "123" = c(significantLSVs_samples_123))

# upset plot
pdf(file="C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/upset_sample_size.pdf", width=3.6, height=3) 
# make upset plot from this information
upset(fromList(significantLSVs_all), 
      order.by = "freq", 
      sets.bar.color = c("#EFC4E4", "#EFC4E4", "#EFC4E4", "#D065B5"), 
      sets.x.label = "# LSVs", 
      mainbar.y.label = "# shared significant LSVs")
dev.off()
```
# Influence on size of detected splicing changes (Fig 4)

Influence on the size of detected splicing changes is determined by plotting the cumulative fraction of dPSIs and probability changing values for each replicate number. 

```{r}
# ecdf for delta psis and probability changing values

ecdf_deltaPSI = ggplot() +
  stat_ecdf(aes(x = maxDeltaPSI), data = sample_sizes_df[[1]], geom = "line", color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "0 1 2"), data = sample_sizes_df[[1]], geom = "line", size = 1, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI), data = sample_sizes_df[[2]], geom = "line", color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "0 1 3"), data = sample_sizes_df[[2]], geom = "line", size = 1, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI), data = sample_sizes_df[[3]], geom = "line", color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "0 2 3"), data = sample_sizes_df[[3]], geom = "line", size = 1, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI), data = sample_sizes_df[[4]], geom = "line", color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxDeltaPSI, color = "1 2 3"), data = sample_sizes_df[[4]], geom = "line", size = 1, pad=FALSE) +
  scale_color_manual(values = c("0 1 2" = "#EFC4E4", "0 1 3" = "#EFC4E4", "0 2 3" = "#EFC4E4", "1 2 3" = "#C76EB1"), 
                     name = "samples", 
                     labels = c("1 2", "1 3", "2 3", "1 2 3")) + 
  coord_cartesian(xlim=c(0, 0.5)) +
  geom_vline(xintercept = 0.05, col = "red3", linetype = "dashed", size = 0.5) +
  ylab("ecdf") +
  xlab("dPSI") +
  ggtitle("C") +
  theme(legend.position="none")

ecdf_probChanging = ggplot() +
  stat_ecdf(aes(x = maxProb), data = sample_sizes_df[[1]], geom = "line", color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxProb, color = "0 1 2"), data = sample_sizes_df[[1]], size = 1, geom = "line", pad=FALSE) +
  
  stat_ecdf(aes(x = maxProb), data = sample_sizes_df[[2]], color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxProb, color = "0 1 3"), data = sample_sizes_df[[2]], size = 1, pad=FALSE) +
  
  stat_ecdf(aes(x = maxProb), data = sample_sizes_df[[3]], color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxProb, color = "0 2 3"), data = sample_sizes_df[[3]], size = 1, geom = "line", pad=FALSE) +
  
  stat_ecdf(aes(x = maxProb), data = sample_sizes_df[[4]], color="black", size = 1.25, pad=FALSE) +
  stat_ecdf(aes(x = maxProb, color = "1 2 3"), data = sample_sizes_df[[4]], size = 1, geom = "line", pad=FALSE) +
  
  scale_color_manual(values = c("0 1 2" = "#EFC4E4", "0 1 3" = "#EFC4E4", "0 2 3" = "#EFC4E4", "1 2 3" = "#C76EB1"), 
                     name = "samples", 
                     labels = c("1 2", "1 3", "2 3", "1 2 3")) + 
  geom_vline(xintercept = 0.05, col = "red3", linetype = "dashed", size = 0.5) +
  ylab("ecdf") +
  xlab("probability changing") +
  ggtitle("C")  +
  theme(legend.position="none")


ecdf_deltaPSI
ecdf_probChanging

# save ggplots
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/ecdf_dPSI_sampleSize.pdf", plot = ecdf_deltaPSI, width = 4.2, height = 4.2, units = "cm")

ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/ecdf_probChanging_sampleSize.pdf", plot = ecdf_probChanging, width = 4.2, height = 4.2, units = "cm")
```

```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE,fig.keep='all'}
sample_sizes_df[[1]]$samples = "0 1 2"
sample_sizes_df[[2]]$samples = "0 1 3"
sample_sizes_df[[3]]$samples = "0 2 3"
sample_sizes_df[[4]]$samples = "1 2 3"

# histograms of dPSIs for each sample combination
ggarrange(
  ggplot(data=sample_sizes_df[[1]], aes(x=maxDeltaPSI, y = ..density..)) +
    geom_histogram(binwidth=.025, fill="#EFC4E4", color="black") +
    #coord_cartesian(xlim = c(0.05, 0.3), ylim = c(0, 5)) +
    #coord_cartesian(xlim = c(0, 0.3), ylim = c(0, 40000)) +
    xlab("\u0394\u03A8") +
    ylab("fraction [%]") +
    ggtitle("sample 1 2"),
  
  ggplot(data=sample_sizes_df[[2]], aes(x=maxDeltaPSI, y = ..density..)) +
    geom_histogram(binwidth=.025, fill="#EFC4E4", color="black") +
    #coord_cartesian(xlim = c(0.05, 0.3), ylim = c(0, 5)) +
    #coord_cartesian(xlim=c(0, 0.3), ylim = c(0, 40000)) +
    xlab("\u0394\u03A8") +
    ylab("fraction [%]") +
    ggtitle("sample 1 3"),
  
  ggplot(data=sample_sizes_df[[3]], aes(x=maxDeltaPSI, y = ..density..)) +
    geom_histogram(binwidth=.025, fill="#EFC4E4", color="black") +
    #coord_cartesian(xlim = c(0.05, 0.3), ylim = c(0, 5)) +
    #coord_cartesian(xlim=c(0, 0.3), ylim = c(0, 40000)) +
    xlab("\u0394\u03A8") +
    ylab("fraction [%]") +
    ggtitle("sample 2 3"),
  
  ggplot(data=sample_sizes_df[[4]], aes(x=maxDeltaPSI, y = ..density..)) +
    geom_histogram(binwidth=.025, fill="#C76EB1", color="black") +
    #coord_cartesian(xlim = c(0.05, 0.3), ylim = c(0, 5)) +
    #coord_cartesian(xlim=c(0, 0.3), ylim = c(0, 40000)) +
    xlab("\u0394\u03A8") +
    ylab("fraction [%]") +
    ggtitle("sample 1 2 3"),
  
  ncol=2, nrow=2
)

# all histograms in one
ggplot(data=NULL, aes(x=maxDeltaPSI, y = ..density..)) +
    geom_histogram(data=sample_sizes_df[[1]], binwidth=.025, fill="#EFC4E4", color="black", alpha=0.5) +
    geom_histogram(data=sample_sizes_df[[2]], binwidth=.025, fill="#EFC4E4", color="black", alpha=0.5) +
    geom_histogram(data=sample_sizes_df[[3]], binwidth=.025, fill="#EFC4E4", color="black", alpha=0.5) +
    geom_histogram(data=sample_sizes_df[[4]], binwidth=.025, fill="blue", color="black", alpha=0.5) +
    coord_cartesian(xlim = c(0.2, 0.5), ylim = c(0, 0.1)) +
    xlab("\u0394\u03A8") +
    ylab("fraction [%]") +
    ggtitle("normalised histogram for all sample combinations") 
```

## Shared binary regulation between the different datasets

Only binary LSVs are considered to ensure that the dPSI value of the same junction is taken from both data sets. 

```{r, warning=FALSE,message=FALSE,error=FALSE,fig.keep='all'}
#################
# all binary LSVs
#################

# filter the data frames for significant LSVs with 2 Junctions
filtered_samples_12 = sample_sizes_df[[1]][sample_sizes_df[[1]]$nJunctions == 2,]

filtered_samples_13 = sample_sizes_df[[2]][sample_sizes_df[[2]]$nJunctions == 2,]

filtered_samples_23 = sample_sizes_df[[3]][sample_sizes_df[[3]]$nJunctions == 2,]

filtered_samples_123 = sample_sizes_df[[4]][sample_sizes_df[[4]]$nJunctions == 2,]


# identify shared significant LSVs for each comparison
# comparison 100/75, 100/50, 100/36
shared_samples_12_13 = intersect(filtered_samples_12$lsv_id, filtered_samples_13$lsv_id) 
length(shared_samples_12_13)

shared_samples_12_23 = intersect(filtered_samples_12$lsv_id, filtered_samples_23$lsv_id) 
length(shared_samples_12_23)

shared_samples_12_123 = intersect(filtered_samples_12$lsv_id, filtered_samples_123$lsv_id) 
length(shared_samples_12_123)

shared_samples_13_23 = intersect(filtered_samples_13$lsv_id, filtered_samples_23$lsv_id) 
length(shared_samples_13_23)

shared_samples_13_123 = intersect(filtered_samples_13$lsv_id, filtered_samples_123$lsv_id) 
length(shared_samples_13_123)

shared_samples_23_123 = intersect(filtered_samples_23$lsv_id, filtered_samples_123$lsv_id) 
length(shared_samples_23_123)


# extract filtered LSVs from the data frames above
# compare samples 1 & 2 with 1 & 3
match_samples_12_13 <- filtered_samples_12[match(shared_samples_12_13, filtered_samples_12$lsv_id),]

match_samples_13_12 <- filtered_samples_13[match(shared_samples_12_13, filtered_samples_13$lsv_id),]

# compare samples 1 & 2 with 2 & 3
match_samples_12_23 <- filtered_samples_12[match(shared_samples_12_23, filtered_samples_12$lsv_id),]

match_samples_23_12 <- filtered_samples_23[match(shared_samples_12_23, filtered_samples_23$lsv_id),]

# compare samples 1 & 2 with 1 & 2 & 3
match_samples_12_123 <- filtered_samples_12[match(shared_samples_12_123, filtered_samples_12$lsv_id),]

match_samples_123_12 <- filtered_samples_123[match(shared_samples_12_123, filtered_samples_123$lsv_id),]

# compare samples 1 & 3 with 2 & 3
match_samples_13_23 <- filtered_samples_13[match(shared_samples_13_23, filtered_samples_13$lsv_id),]

match_samples_23_13 <- filtered_samples_23[match(shared_samples_13_23, filtered_samples_23$lsv_id),]

# compare samples 1 & 3 with 1 & 2 & 3
match_samples_13_123 <- filtered_samples_13[match(shared_samples_13_123, filtered_samples_13$lsv_id),]

match_samples_123_13 <- filtered_samples_123[match(shared_samples_13_123, filtered_samples_123$lsv_id),]

# compare samples 2 & 3 with 1 & 2 & 3
match_samples_23_123 <- filtered_samples_23[match(shared_samples_23_123, filtered_samples_23$lsv_id),]

match_samples_123_23 <- filtered_samples_123[match(shared_samples_23_123, filtered_samples_123$lsv_id),]


# extract dpsi value of the first junction for all LSVs
# deltapsi samples 1 & 2 vs samples 1 & 3
deltaPSI_samples_12_vs_13 = match_samples_12_13$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_13_vs_12 = match_samples_13_12$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 2 vs samples 2 & 3
deltaPSI_samples_12_vs_23 = match_samples_12_23$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_23_vs_12 = match_samples_23_12$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 2 vs samples 1 & 2 & 3
deltaPSI_samples_12_vs_123 = match_samples_12_123$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_123_vs_12 = match_samples_123_12$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 3 vs samples 2 & 3
deltaPSI_samples_13_vs_23 = match_samples_13_23$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_23_vs_13 = match_samples_23_13$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 3 vs samples 1 & 2 & 3
deltaPSI_samples_13_vs_123 = match_samples_13_123$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_123_vs_13 = match_samples_123_13$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 2 & 3 vs samples 1 & 2 & 3
deltaPSI_samples_23_vs_123 = match_samples_23_123$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_123_vs_23 = match_samples_123_23$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric


#########################
# significant binary LSVs
#########################

# filter the data frames for significant LSVs with 2 Junctions
filtered_samples_12_sig = sample_sizes_df[[1]][sample_sizes_df[[1]]$nJunctions == 2 & sample_sizes_df[[1]]$significant == TRUE,]

filtered_samples_13_sig = sample_sizes_df[[2]][sample_sizes_df[[2]]$nJunctions == 2 & sample_sizes_df[[2]]$significant == TRUE,]

filtered_samples_23_sig = sample_sizes_df[[3]][sample_sizes_df[[3]]$nJunctions == 2 & sample_sizes_df[[3]]$significant == TRUE,]

filtered_samples_123_sig = sample_sizes_df[[4]][sample_sizes_df[[4]]$nJunctions == 2 & sample_sizes_df[[4]]$significant == TRUE,]


# identify shared significant LSVs for each comparison
# comparison 100/75, 100/50, 100/36
shared_samples_12_13_sig = intersect(filtered_samples_12_sig$lsv_id, filtered_samples_13_sig$lsv_id) 
length(shared_samples_12_13_sig)

shared_samples_12_23_sig = intersect(filtered_samples_12_sig$lsv_id, filtered_samples_23_sig$lsv_id) 
length(shared_samples_12_23_sig)

shared_samples_12_123_sig = intersect(filtered_samples_12_sig$lsv_id, filtered_samples_123_sig$lsv_id) 
length(shared_samples_12_123_sig)

shared_samples_13_23_sig = intersect(filtered_samples_13_sig$lsv_id, filtered_samples_23_sig$lsv_id) 
length(shared_samples_13_23_sig)

shared_samples_13_123_sig = intersect(filtered_samples_13_sig$lsv_id, filtered_samples_123_sig$lsv_id) 
length(shared_samples_13_123_sig)

shared_samples_23_123_sig = intersect(filtered_samples_23_sig$lsv_id, filtered_samples_123_sig$lsv_id) 
length(shared_samples_23_123_sig)


# extract filtered LSVs from the data frames above
# compare samples 1 & 2 with 1 & 3
match_samples_12_13_sig<- filtered_samples_12_sig[match(shared_samples_12_13_sig, filtered_samples_12_sig$lsv_id),]

match_samples_13_12_sig<- filtered_samples_13_sig[match(shared_samples_12_13_sig, filtered_samples_13_sig$lsv_id),]

# compare samples 1 & 2 with 2 & 3
match_samples_12_23_sig<- filtered_samples_12_sig[match(shared_samples_12_23_sig, filtered_samples_12_sig$lsv_id),]

match_samples_23_12_sig<- filtered_samples_23_sig[match(shared_samples_12_23_sig, filtered_samples_23_sig$lsv_id),]

# compare samples 1 & 2 with 1 & 2 & 3
match_samples_12_123_sig<- filtered_samples_12_sig[match(shared_samples_12_123_sig, filtered_samples_12_sig$lsv_id),]

match_samples_123_12_sig<- filtered_samples_123_sig[match(shared_samples_12_123_sig, filtered_samples_123_sig$lsv_id),]

# compare samples 1 & 3_sigwith 2 & 3
match_samples_13_23_sig<- filtered_samples_13_sig[match(shared_samples_13_23_sig, filtered_samples_13_sig$lsv_id),]

match_samples_23_13_sig<- filtered_samples_23_sig[match(shared_samples_13_23_sig, filtered_samples_23_sig$lsv_id),]

# compare samples 1 & 3_sigwith 1 & 2 & 3
match_samples_13_123_sig<- filtered_samples_13_sig[match(shared_samples_13_123_sig, filtered_samples_13_sig$lsv_id),]

match_samples_123_13_sig<- filtered_samples_123_sig[match(shared_samples_13_123_sig, filtered_samples_123_sig$lsv_id),]

# compare samples 2 & 3_sigwith 1 & 2 & 3
match_samples_23_123_sig<- filtered_samples_23_sig[match(shared_samples_23_123_sig, filtered_samples_23_sig$lsv_id),]

match_samples_123_23_sig<- filtered_samples_123_sig[match(shared_samples_23_123_sig, filtered_samples_123_sig$lsv_id),]


# extract dpsi value of the first junction for all LSVs
# deltapsi samples 1 & 2 vs samples 1 & 3
deltaPSI_samples_12_vs_13_sig = match_samples_12_13_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_13_vs_12_sig = match_samples_13_12_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 2 vs samples 2 & 3
deltaPSI_samples_12_vs_23_sig = match_samples_12_23_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_23_vs_12_sig = match_samples_23_12_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 2 vs samples 1 & 2 & 3
deltaPSI_samples_12_vs_123_sig = match_samples_12_123_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_123_vs_12_sig = match_samples_123_12_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 3 vs samples 2 & 3
deltaPSI_samples_13_vs_23_sig = match_samples_13_23_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_23_vs_13_sig = match_samples_23_13_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 1 & 3 vs samples 1 & 2 & 3
deltaPSI_samples_13_vs_123_sig = match_samples_13_123_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_123_vs_13_sig = match_samples_123_13_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

# deltapsi samples 2 & 3 vs samples 1 & 2 & 3
deltaPSI_samples_23_vs_123_sig = match_samples_23_123_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric

deltaPSI_samples_123_vs_23_sig = match_samples_123_23_sig$mean_dpsi_per_lsv_junction %>%
  strsplit(split = ";") %>% 
  sapply(FUN = "[[",1) %>% as.numeric
```

```{r, warning=FALSE,message=FALSE,error=FALSE,fig.keep='all'}
# build data frame out of vectors with dpsi values 
# create scatter plot with the data frames

# correlation data frame samples 12 vs 13
correlationDataFrame_12_vs_13 <- data.frame(deltaPSI_samples_12_vs_13, deltaPSI_samples_13_vs_12)
correlationDataFrame_12_vs_13_sig <- data.frame(deltaPSI_samples_12_vs_13_sig, deltaPSI_samples_13_vs_12_sig)

sharedRegulation_scatterPlot1 = ggplot(data=correlationDataFrame_12_vs_13, aes(x=deltaPSI_samples_12_vs_13, deltaPSI_samples_13_vs_12)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey") +
  geom_smooth(data=correlationDataFrame_12_vs_13, aes(x=deltaPSI_samples_12_vs_13, deltaPSI_samples_13_vs_12), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_12_vs_13_sig, aes(x=deltaPSI_samples_12_vs_13_sig, y=deltaPSI_samples_13_vs_12_sig), color="black") +
  geom_smooth(data=correlationDataFrame_12_vs_13_sig, aes(x=deltaPSI_samples_12_vs_13_sig, y=deltaPSI_samples_13_vs_12_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("\u0394\u03A8 samples 1 & 2") +
  ylab("\u0394\u03A8 samples 1 & 3") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame samples 12 vs 23
correlationDataFrame_12_vs_23 <- data.frame(deltaPSI_samples_12_vs_23, deltaPSI_samples_23_vs_12)
correlationDataFrame_12_vs_23_sig <- data.frame(deltaPSI_samples_12_vs_23_sig, deltaPSI_samples_23_vs_12_sig)

sharedRegulation_scatterPlot2 = ggplot(data=correlationDataFrame_12_vs_23, aes(x=deltaPSI_samples_12_vs_23, deltaPSI_samples_23_vs_12)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey") +
  geom_smooth(data=correlationDataFrame_12_vs_23, aes(x=deltaPSI_samples_12_vs_23, deltaPSI_samples_23_vs_12), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_12_vs_23_sig, aes(x=deltaPSI_samples_12_vs_23_sig, y=deltaPSI_samples_23_vs_12_sig), color="black") +
  geom_smooth(data=correlationDataFrame_12_vs_23_sig, aes(x=deltaPSI_samples_12_vs_23_sig, y=deltaPSI_samples_23_vs_12_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("\u0394\u03A8 samples 1 & 2") +
  ylab("\u0394\u03A8 samples 2 & 3") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame samples 12 vs 123
correlationDataFrame_12_vs_123 <- data.frame(deltaPSI_samples_12_vs_123, deltaPSI_samples_123_vs_12)
correlationDataFrame_12_vs_123_sig <- data.frame(deltaPSI_samples_12_vs_123_sig, deltaPSI_samples_123_vs_12_sig)

sharedRegulation_scatterPlot3 = ggplot(data=correlationDataFrame_12_vs_123, aes(x=deltaPSI_samples_12_vs_123, deltaPSI_samples_123_vs_12)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey", size=0.3) +
  geom_smooth(data=correlationDataFrame_12_vs_123, aes(x=deltaPSI_samples_12_vs_123, deltaPSI_samples_123_vs_12), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_12_vs_123_sig, aes(x=deltaPSI_samples_12_vs_123_sig, y=deltaPSI_samples_123_vs_12_sig), color="black", size=0.3) +
  geom_smooth(data=correlationDataFrame_12_vs_123_sig, aes(x=deltaPSI_samples_12_vs_123_sig, y=deltaPSI_samples_123_vs_12_sig),method = "lm", color="black") +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1)) +
  xlab("dPSI samples 1 & 2") +
  ylab("dPSI samples 1 & 2 & 3") +
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(plot.title = element_text(face = "bold")) +
  ggtitle("F")

# correlation data frame samples 13 vs 23
correlationDataFrame_13_vs_23 <- data.frame(deltaPSI_samples_13_vs_23, deltaPSI_samples_23_vs_13)
correlationDataFrame_13_vs_23_sig <- data.frame(deltaPSI_samples_13_vs_23_sig, deltaPSI_samples_23_vs_13_sig)

sharedRegulation_scatterPlot4 = ggplot(data=correlationDataFrame_13_vs_23, aes(x=deltaPSI_samples_13_vs_23, deltaPSI_samples_23_vs_13)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey") +
  geom_smooth(data=correlationDataFrame_13_vs_23, aes(x=deltaPSI_samples_13_vs_23, deltaPSI_samples_23_vs_13), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_13_vs_23_sig, aes(x=deltaPSI_samples_13_vs_23_sig, y=deltaPSI_samples_23_vs_13_sig), color="black") +
  geom_smooth(data=correlationDataFrame_13_vs_23_sig, aes(x=deltaPSI_samples_13_vs_23_sig, y=deltaPSI_samples_23_vs_13_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("\u0394\u03A8 samples 1 & 3") +
  ylab("\u0394\u03A8 samples 2 & 3") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame samples 13 vs 123
correlationDataFrame_13_vs_123 <- data.frame(deltaPSI_samples_13_vs_123, deltaPSI_samples_123_vs_13)
correlationDataFrame_13_vs_123_sig <- data.frame(deltaPSI_samples_13_vs_123_sig, deltaPSI_samples_123_vs_13_sig)

sharedRegulation_scatterPlot5 = ggplot(data=correlationDataFrame_13_vs_123, aes(x=deltaPSI_samples_13_vs_123, deltaPSI_samples_123_vs_13)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey") +
  geom_smooth(data=correlationDataFrame_13_vs_123, aes(x=deltaPSI_samples_13_vs_123, deltaPSI_samples_123_vs_13), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_13_vs_123_sig, aes(x=deltaPSI_samples_13_vs_123_sig, y=deltaPSI_samples_123_vs_13_sig), color="black") +
  geom_smooth(data=correlationDataFrame_13_vs_123_sig, aes(x=deltaPSI_samples_13_vs_123_sig, y=deltaPSI_samples_123_vs_13_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("\u0394\u03A8 samples 1 & 3") +
  ylab("\u0394\u03A8 samples 1 & 2 & 3") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# correlation data frame samples 23 vs 123
correlationDataFrame_23_vs_123 <- data.frame(deltaPSI_samples_23_vs_123, deltaPSI_samples_123_vs_23)
correlationDataFrame_23_vs_123_sig <- data.frame(deltaPSI_samples_23_vs_123_sig, deltaPSI_samples_123_vs_23_sig)

sharedRegulation_scatterPlot6 = ggplot(data=correlationDataFrame_23_vs_123, aes(x=deltaPSI_samples_23_vs_123, deltaPSI_samples_123_vs_23)) +
  stat_cor(label.x.npc = .3, label.y.npc =1) +
  ggrastr::geom_point_rast(alpha=0.36, color="darkgrey") +
  geom_smooth(data=correlationDataFrame_23_vs_123, aes(x=deltaPSI_samples_23_vs_123, deltaPSI_samples_123_vs_23), method = "lm", color="darkgrey", se=FALSE) +
  ggrastr::geom_point_rast(data=correlationDataFrame_23_vs_123_sig, aes(x=deltaPSI_samples_23_vs_123_sig, y=deltaPSI_samples_123_vs_23_sig), color="black") +
  geom_smooth(data=correlationDataFrame_23_vs_123_sig, aes(x=deltaPSI_samples_23_vs_123_sig, y=deltaPSI_samples_123_vs_23_sig),method = "lm", color="black", se=FALSE) +
  geom_hline(yintercept=0, size=.5, linetype="dashed") +
  geom_vline(xintercept=0, size=.5, linetype="dashed") +
  geom_abline(xintercept=0, slope=1, linetype="dashed", size=.75, color="firebrick2") +
  coord_fixed(ratio = 1) +
  xlab("\u0394\u03A8 samples 2 & 3") +
  ylab("\u0394\u03A8 samples 1 & 2 & 3") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

sharedRegulation_scatterPlot1
sharedRegulation_scatterPlot2
sharedRegulation_scatterPlot3
sharedRegulation_scatterPlot4
sharedRegulation_scatterPlot5
sharedRegulation_scatterPlot6

'# save ggplots
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/sharedRehulation_100_75.pdf", plot = U2AF2_sharedRegulation_scatterPlot1)
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/sharedRehulation_100_50.pdf", plot = U2AF2_sharedRegulation_scatterPlot2)
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/sharedRehulation_100_36.pdf", plot = U2AF2_sharedRegulation_scatterPlot3)
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/sharedRehulation_75_50.pdf", plot = U2AF2_sharedRegulation_scatterPlot4)
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/sharedRehulation_50_36.pdf", plot = U2AF2_sharedRegulation_scatterPlot5)'

ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/sharedRegulation_samples_12_vs_123_sampleSize.pdf", plot = sharedRegulation_scatterPlot3, width = 5.65, height = 5.65, units = "cm")
```


# 5 - Influence on the distribution of alternative splice events types (Fig 5)

Since there are alternative splicing events with different numbers of junctions, they are treated differently in terms of significance. 

Alternative splicing events with two junctions (A3SS, A5SS, AFE, ALE, IR) are considered significantly regulated if at least one junction has a dPSI of at least 0.05 and a probability changing value of at least 0.9. Events with 4 junctions (CE and MXE) are considered significantly regulated if either both junctions starting from C1 or both junctions starting from C2 fulfill the regulation criteria. 

```{r}
# load voila modulize files 
modDir = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/endodermal_cell_vs_H1_cell_line/Modulize"
modFiles = list.files(path = modDir, pattern = ".tsv")
setwd(modDir)
modFiles

# list events with 2 Junctions separately for each replicate number 
events_2J = lapply(modFiles[1:20], loadFiles)
events_4J = lapply(modFiles[21:28], loadFiles)

# pseudo column for events with 2 junctions
# to ensure the same structure of the data frames
for (i in 9:16){
  events_2J[[i]] <- data.frame(pseudoCol = NA, events_2J[[i]])
}

# pseudo column for events with 4 junctions
for (i in 5:8){
  events_4J[[i]] <- data.frame(pseudoCol = NA, events_4J[[i]])
}

# function to filter out significant events with 2 junctions
# significant events (dpsi >= 0.05, probability changing >= 0.9)
filter2JEvents <- function(eventTable) {
  names(eventTable)[21] <- "CT.KD_median_dpsi"
  names(eventTable)[22] <- "CT.KD_probability_changing"
  eventTable = eventTable[abs(as.numeric(eventTable$CT.KD_median_dpsi)) >= 0.05 & eventTable$CT.KD_probability_changing >= 0.9, ] 
  eventTable %>% group_by(event_id) %>% slice(1)
  }

# function to filter out significant events with 4 junctions
filter4JEvents <- function(eventTable) {
  names(eventTable)[21] <- "CT.KD_median_dpsi"
  names(eventTable)[22] <- "CT.KD_probability_changing"
  eventTable %>% 
    group_by(event_id) %>%
    summarise(significant = (((abs(as.numeric(CT.KD_median_dpsi[1])) >= 0.05 & CT.KD_probability_changing[1] >= 0.9) &
    (abs(as.numeric(CT.KD_median_dpsi[2])) >= 0.05 & CT.KD_probability_changing[2] >= 0.9)) |
  ((abs(as.numeric(CT.KD_median_dpsi[3])) >= 0.05 & CT.KD_probability_changing[3] >= 0.9) &
    (abs(as.numeric(CT.KD_median_dpsi[4])) >= 0.05 & CT.KD_probability_changing[4] >= 0.9)))) %>%
    filter(significant == TRUE)
}

# filter out significant events
significantEvents_2J = lapply(events_2J, filter2JEvents)
significantEvents_4J = lapply(events_4J, filter4JEvents)
names(significantEvents_2J) = c("A3SS_1_2", "A3SS_1_3", "A3SS_2_3", "A3SS_1_2_3", 
                                "A5SS_1_2", "A5SS_1_3", "A5SS_2_3", "A5SS_1_2_3", 
                                "AFE_1_2", "AFE_1_3", "AFE_2_3", "AFE_1_2_3", 
                                "ALE_1_2", "ALE_1_3", "ALE_2_3", "ALE_1_2_3", 
                                "IR_1_2", "IR_1_3", "IR_2_3", "IR_1_2_3")
names(significantEvents_4J) = c("CE_1_2", "CE_1_3", "CE_2_3", "CE_1_2_3",
                                "MXE_1_2", "MXE_1_3", "MXE_2_3", "MXE_1_2_3")

# data frames with number of event types for each replicate number
nEventTypes_sample_1_2_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE", "MXE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_1_2), 
                                             nrow(significantEvents_2J$A5SS_1_2),
                                             nrow(significantEvents_2J$AFE_1_2), 
                                             nrow(significantEvents_2J$ALE_1_2), 
                                             nrow(significantEvents_2J$IR_1_2), 
                                             nrow(significantEvents_4J$CE_1_2),
                                             nrow(significantEvents_4J$MXE_1_2)))

nEventTypes_sample_1_3_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE", "MXE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_1_3), 
                                             nrow(significantEvents_2J$A5SS_1_3),
                                             nrow(significantEvents_2J$AFE_1_3), 
                                             nrow(significantEvents_2J$ALE_1_3), 
                                             nrow(significantEvents_2J$IR_1_3), 
                                             nrow(significantEvents_4J$CE_1_3),
                                             nrow(significantEvents_4J$MXE_1_3)))

nEventTypes_sample_2_3_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE", "MXE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_2_3), 
                                             nrow(significantEvents_2J$A5SS_2_3),
                                             nrow(significantEvents_2J$AFE_2_3), 
                                             nrow(significantEvents_2J$ALE_2_3), 
                                             nrow(significantEvents_2J$IR_2_3), 
                                             nrow(significantEvents_4J$CE_2_3),
                                             nrow(significantEvents_4J$MXE_2_3)))

nEventTypes_sample_1_2_3_df = data.frame(event = c("A3SS", "A5SS", "AFE", "ALE", "IR", "CE", "MXE"), 
                                       n = c(nrow(significantEvents_2J$A3SS_1_2_3), 
                                             nrow(significantEvents_2J$A5SS_1_2_3),
                                             nrow(significantEvents_2J$AFE_1_2_3), 
                                             nrow(significantEvents_2J$ALE_1_2_3), 
                                             nrow(significantEvents_2J$IR_1_2_3), 
                                             nrow(significantEvents_4J$CE_1_2_3),
                                             nrow(significantEvents_4J$MXE_1_2_3)))


# combine event type results of each replicate number into one data frame
nEventTypes_sample_1_2_df$samples = "012"
nEventTypes_sample_1_3_df$samples = "013"
nEventTypes_sample_2_3_df$samples = "023"
nEventTypes_sample_1_2_3_df$samples = "123"
nEventTypes_df = rbind(nEventTypes_sample_1_2_df, 
                         nEventTypes_sample_1_3_df, 
                         nEventTypes_sample_2_3_df, 
                         nEventTypes_sample_1_2_3_df)

# number of significant event types for each replicate number as bar charts
nEventTypes_cols = 
  ggplot(nEventTypes_df, aes(x = event, y = n, fill = samples)) +
  geom_col(position = "dodge") +
  xlab("event type") +
  ylab("# events") +
  scale_fill_manual(values = c("012" = "#EFC4E4", "013" = "#EFC4E4", "023" = "#EFC4E4", "123" = "#C76EB1"), 
                    name = "# samples", 
                    labels = c("12", "13", "23", "123"))
 
# fraction of significant event types for each replicate number as stacked bar chart 
# amount of events log10 transformed
nEventTypes_StackedBarChart = 
  ggplot(nEventTypes_df, aes(x = samples, y = n, fill = event)) +
  geom_col(position = "fill") +
  scale_y_log10() +
  xlab("# samples") +
  ylab("fraction of event types (log10)") +
  scale_fill_brewer(palette = "YlGnBu", name = "Event") +
  ggtitle("C")

nEventTypes_cols
nEventTypes_StackedBarChart

# save plots 
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/nEventTypes_SS.pdf", plot = nEventTypes_cols, width = 11, height = 7, units = "cm")
ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/distrEventTypes_SS.pdf", plot = nEventTypes_StackedBarChart, width = 4.8, height = 7, units = "cm")

# wieso sind hier zwei NA Werte? wie kommen die zustande?
# entsprechende Zeilen ausprinten lassen
'print(sum(is.na(filteredEvents_2J[[20]]$significant)))
rows_with_na <- subset(filteredEvents_2J[[20]], is.na(significant))
print(rows_with_na)'
```

```{r}
#################################
# fraction of significant events
#################################

# pseudo column for AFE events with 2 junctions
# to ensure the same structure of the data frames
'for (i in 9:16){
  events_2J[[i]] <- data.frame(pseudoCol = NA, events_2J[[i]])
}

# pseudo column for MXE events with 4 junctions
for (i in 5:8){
  events_4J[[i]] <- data.frame(pseudoCol = NA, events_4J[[i]])
}'

# function to filter significant events with 2 junctions
# significant events (dpsi >= 0.05, probability changing >= 0.9)
filter2JEvents <- function(eventTable) {
  names(eventTable)[21] = "CT.KD_median_dpsi"
  names(eventTable)[22] = "CT.KD_probability_changing"
  # add new column "significant" for each replicate number which could be TRUE/FALSE
  eventTable$significant = abs(as.numeric(eventTable$CT.KD_median_dpsi)) >= 0.05 & 
    eventTable$CT.KD_probability_changing >= 0.9 
  # select only one entry for each event id (if for an entry significant = TRUE, it will be taken)
  eventTable = eventTable %>%
    group_by(event_id) %>%
    # sort entries in descending order to ensure that entries with significant = TRUE are selected in any case
    arrange(desc((significant))) %>%
    slice(1)
}

# function to filter out significant events with 4 junctions
filter4JEvents <- function(eventTable) {
  names(eventTable)[21] <- "CT.KD_median_dpsi"
  names(eventTable)[22] <- "CT.KD_probability_changing"
  eventTable$significant = (((abs(as.numeric(eventTable$CT.KD_median_dpsi)) >= 0.05 & eventTable$CT.KD_probability_changing >= 0.9) &
    (abs(as.numeric(eventTable$CT.KD_median_dpsi)) >= 0.05 & eventTable$CT.KD_probability_changing >= 0.9)) |
  ((abs(as.numeric(eventTable$CT.KD_median_dpsi)) >= 0.05 & eventTable$CT.KD_probability_changing >= 0.9) &
    (abs(as.numeric(eventTable$CT.KD_median_dpsi)) >= 0.05 & eventTable$CT.KD_probability_changing >= 0.9)))
  eventTable = eventTable %>% 
    group_by(event_id) %>%
    arrange(desc((significant))) %>%
    slice(1)
}

# filter significant events
filteredEvents_2J = lapply(events_2J, filter2JEvents)
filteredEvents_4J = lapply(events_4J, filter4JEvents)

# significant events for each sample size
print("# significant events with 2 junctions")
for (i in 1:length(filteredEvents_2J)) {
  print(i)
  print(table(filteredEvents_2J[[i]]$significant))
}

print("# significant events with 4 junctions")
for (i in 1:length(filteredEvents_4J)) {
  print(i)
  print(table(filteredEvents_4J[[i]]$significant))
}

# test if all data frames have the same number of columns
for (i in 1:length(filteredEvents_2J)) {
  print(ncol(filteredEvents_2J[[i]]))
}
for (i in 1:length(filteredEvents_4J)) {
  print(ncol(filteredEvents_4J[[i]]))
}
```

Immer aufschreiben, wie ich die Plots gemacht habe, die in den Figures sind! Damit ich es spter noch nachvollziehen kann. 

Fragen:
- Wieso habe ich jetzt nur die IRs geplottet (bzw. generell die ganzen Eventtypen einzeln), die fraction?? Habe ich diese manuell zusammengebastelt? Wahrscheinlich ja. Wie bin ich dann auf den Durchschnitt fr die 2 samples gekommen? In Affinity designer gemacht? Wahrscheinlich auch ja.
Antwort: Habe Plots auf PC unter plots gefunden. Sind doch nicht einzeln, sondern alle events zsm in einem Plot. Wie habe ich das gemacht? Code ist warum auch immer weg. 

This plot was generated individually for all event types.
```{r}
# create data frame with a column for the sample combination
# and a column indicating whether the associated event is significant or not 
significantEvents_df = data.frame(comparison=c(rep("012", nrow(filteredEvents_4J[[5]])),
                                                   rep("013", nrow(filteredEvents_4J[[6]])), 
                                                   rep("023", nrow(filteredEvents_4J[[7]])), 
                                                   rep("123", nrow(filteredEvents_4J[[8]]))),
                                      significant=c(filteredEvents_4J[[5]]$significant,
                                                    filteredEvents_4J[[6]]$significant, 
                                                    filteredEvents_4J[[7]]$significant, 
                                                    filteredEvents_4J[[8]]$significant))

# differ between the significance of events of different sample combinations to plot them in different colors 
for (i in 1:nrow(significantEvents_df)){
  if (significantEvents_df[i, "comparison"] == "012" && significantEvents_df[i, "significant"] == TRUE){
    significantEvents_df[i, "significant"] = "TRUE replicates 012"
  }
  if (significantEvents_df[i, "comparison"] == "013" && significantEvents_df[i, "significant"] == TRUE){
    significantEvents_df[i, "significant"] = "TRUE replicates 013"
  }
  if (significantEvents_df[i, "comparison"] == "023" && significantEvents_df[i, "significant"] == TRUE){
    significantEvents_df[i, "significant"] = "TRUE replicates 023"
  }
  if (significantEvents_df[i, "comparison"] == "123" && significantEvents_df[i, "significant"] == TRUE){
    significantEvents_df[i, "significant"] = "TRUE replicates 123"
  }
}

# stacked bar chart showing fraction of significant events among all detected events for each sample size   
significantEvents_fraction = ggplot(data=significantEvents_df, aes(x = comparison, fill = factor(significant))) +
  geom_bar(position = "fill", show.legend = FALSE) +
  xlab("event type") +
  ylab("fraction of significant detected events [%]") +
  coord_cartesian(ylim = c(0, 0.15)) +
  scale_fill_manual(name= "significant", values = c("lightgrey", "#EFC4E4", "#EFC4E4", "#EFC4E4", "#C76EB1"), labels = c("FALSE", "TRUE", "TRUE", "TRUE", "TRUE")) +
  scale_x_discrete(labels = c("12", "13", "23", "123")) +
  ggtitle("MXE")

significantEvents_fraction
significantEvents_df

filteredEvents_4J[[5]]
```
```{r}
'ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/fractionEventTypes_SS.pdf", plot = significantEvents_fraction, width = 11, height = 7, units = "cm")'
```
There are IR events that have no dpsi values. These must be removed beforehand, as it is not possible to determine whether they are significant or not.
```{r}
# make all fraction bar charts in one plot
# x stands for the mean of the different sample size combinations 
# define event types and replicates
event_types <- c("A3SS", "A5SS", "AFE", "ALE", "CE", "IR", "MXE")
replicates <- c("012", "013", "023", "x" ,"123")

# create df with all combinations of event types and replicates
sig_events_df <- expand.grid(replicates = replicates, event_type = event_types)

# count number of significant and not significant events for each event type and number of replicates 
# also remove rows with NAs
sig_2J <- list()
sig_4J <- list()
not_sig_2J <- list()
not_sig_4J <- list()

for (i in 1:20) {
  sig_2J[[i]] <- sum(filteredEvents_2J[[i]]$significant == TRUE, na.rm = TRUE)
}
for (i in 1:8) {
  sig_4J[[i]] <- sum(filteredEvents_4J[[i]]$significant == TRUE, na.rm = TRUE)
}

for (i in 1:20) {
  not_sig_2J[[i]] <- sum(filteredEvents_2J[[i]]$significant == FALSE, na.rm = TRUE)
}
for (i in 1:8) {
  not_sig_4J[[i]] <- sum(filteredEvents_4J[[i]]$significant == FALSE, na.rm = TRUE)
}

# add information to data frame
# remove rows with NA in column 'significant'
sig_events_df$significant <- c(sig_2J[[1]], sig_2J[[2]], sig_2J[[3]], mean(c(sig_2J[[1]], sig_2J[[2]], sig_2J[[3]])), sig_2J[[4]],
                               sig_2J[[5]], sig_2J[[6]], sig_2J[[7]], mean(c(sig_2J[[5]], sig_2J[[6]], sig_2J[[7]])), sig_2J[[8]],
                               sig_2J[[9]], sig_2J[[10]], sig_2J[[11]], mean(c(sig_2J[[9]], sig_2J[[10]], sig_2J[[11]])), sig_2J[[12]],
                               sig_2J[[13]], sig_2J[[14]], sig_2J[[15]], mean(c(sig_2J[[13]], sig_2J[[14]], sig_2J[[15]])), sig_2J[[16]],
                               sig_4J[[1]], sig_4J[[2]], sig_4J[[3]], mean(c(sig_4J[[1]], sig_4J[[2]], sig_4J[[3]])), sig_4J[[4]],
                               sig_2J[[17]], sig_2J[[18]], sig_2J[[19]], mean(c(sig_2J[[17]], sig_2J[[18]], sig_2J[[19]])), sig_2J[[20]],
                               sig_4J[[5]], sig_4J[[6]], sig_4J[[7]], mean(c(sig_4J[[5]], sig_4J[[6]], sig_4J[[7]])), sig_4J[[8]])

sig_events_df$not_significant <- c(not_sig_2J[[1]], not_sig_2J[[2]], not_sig_2J[[3]], 
                                   mean(c(not_sig_2J[[1]], not_sig_2J[[2]], not_sig_2J[[3]])), not_sig_2J[[4]],
                                   not_sig_2J[[5]], not_sig_2J[[6]], not_sig_2J[[7]], 
                                   mean(c(not_sig_2J[[5]], not_sig_2J[[6]], not_sig_2J[[7]])), not_sig_2J[[8]],
                                   not_sig_2J[[9]], not_sig_2J[[10]], not_sig_2J[[11]], 
                                   mean(c(not_sig_2J[[9]], not_sig_2J[[10]], not_sig_2J[[11]])), not_sig_2J[[12]],
                                   not_sig_2J[[13]], not_sig_2J[[14]], not_sig_2J[[15]], 
                                   mean(c(not_sig_2J[[13]], not_sig_2J[[14]], not_sig_2J[[15]])), not_sig_2J[[16]],
                                   not_sig_4J[[1]], not_sig_4J[[2]], not_sig_4J[[3]], 
                                   mean(c(not_sig_4J[[1]], not_sig_4J[[2]], not_sig_4J[[3]])), not_sig_4J[[4]],
                                   not_sig_2J[[17]], not_sig_2J[[18]], not_sig_2J[[19]], 
                                   mean(c(not_sig_2J[[17]], not_sig_2J[[18]], not_sig_2J[[19]])), not_sig_2J[[20]],
                                   not_sig_4J[[5]], not_sig_4J[[6]], not_sig_4J[[7]], 
                                   mean(c(not_sig_4J[[5]], not_sig_4J[[6]], not_sig_4J[[7]])), not_sig_4J[[8]])
sig_events_df

# Berechnen Sie den Anteil an signifikanten Ereignissen
sig_events_df$percentage <- sig_events_df$significant / (sig_events_df$significant + sig_events_df$not_significant) * 100

# Erstellen Sie ein gestapeltes Balkendiagramm fr jeden Event-Typ
fractionSigEvents = ggplot(sig_events_df, aes(x = replicates, y = percentage, fill = replicates)) +
  geom_bar(stat = "identity") +
  facet_wrap(~event_type, ncol = 7) +
  ggtitle("C") +
  ylab("fraction of significant detected events [%]") +
  xlab("event type") +
  scale_fill_manual(values = c("012" = "#EFC4E4", "013" = "#EFC4E4", "023" = "#EFC4E4", "123" = "#C76EB1")) 

sig_events_df

ggsave(filename = "C:/Users/ladwi/Desktop/HiWi/project_impact_read_length/plots/fractionSigEvents_SS.pdf", plot = fractionSigEvents, width = 13.3, height = 6, units = "cm")
```


# Session info

```{r}
sessionInfo()
```
